---
title: "HFNO Post-Extubation"
subtitle: "Discrete-Time Competing Risks with Marginal Structural Models"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    embed-resources: true
    fig-width: 11
    fig-height: 6
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This document implements the pre-specified  analysis for the HFNO post-extubation study. Sites should run this document from top to bottom without modification. All outputs are saved to the `output_to_share/` directory for return to the coordinating site.

**Study Design:**

- **Exposure:** Continuous HFNO flow rate (L/min), time-varying over 4-hour windows
- **Outcomes:** Competing risks — reintubation/death (failure) vs. discharge alive (success)
- **Method:** Discrete-time marginal structural model with inverse probability of treatment weighting
- **Follow-up:** Up to 7 days (42 windows) post-extubation
- **TV Confounder:** Lagged FiO2 (primary time-varying confounder, pre-specified based on pilot site)

**Required data file:** `hfno_trajectory.parquet` in the working directory

**Required variables:**

| Variable | Description |
|---|---|
| `hospitalization_id` | Unique hospitalization identifier |
| `patient_id` | Unique patient identifier |
| `window_idx` | 4-hour window index (0 = extubation window) |
| `hour_window` | Hour post-extubation |
| `hfno_flow_rate` | HFNO flow rate (L/min) |
| `hfno_fio2` | FiO2 at window |
| `sf_ratio` | S/F ratio at window |
| `rox_index` | ROX index at window |
| `outcome_death_reintubate` | 0=at risk, 1=failure, 2=discharge |
| `outcome_death_reintubate_nippv_cpap` | Secondary outcome (NIPPV/CPAP as failure) |
| `age`, `sex`, `race`, `bmi`, `cci` | Baseline demographics |
| `sofa_extubation` | SOFA score at extubation |
| `is_vaso` | Vasopressor use at window |
| `hospital_id` | Facility identifier (set to constant if single facility) |
| `extubation_icu_type` | ICU identifier (set to constant if single ICU) |

---

# Section 0 — Setup

```{r setup}
library(tidyverse)
library(arrow)
library(splines)
library(gtsummary)
library(broom)
library(patchwork)
library(zoo)
library(nnet)
library(mclogit)
library(knitr)
library(scales)
install.packages("here")
library(here)

set.seed(20240101)

# Create output directories
dir.create(here("output_to_share/figures"),     showWarnings = FALSE, recursive = TRUE)
dir.create(here("output_to_share/tables"),      showWarnings = FALSE, recursive = TRUE)
dir.create(here("output_to_share/diagnostics"), showWarnings = FALSE, recursive = TRUE)
dir.create(here("output_to_share/models"),      showWarnings = FALSE, recursive = TRUE)

# Load data
df_raw <- read_parquet("../output/hfno_trajectory.parquet")

# Outcome color palette (used throughout)
outcome_colors <- c(
  "Reintubation/Death" = "#C0392B",
  "Discharge Alive"    = "#27AE60",
  "Censored"           = "#2980B9"
)

cat("Data loaded:", nrow(df_raw), "rows,", 
    n_distinct(df_raw$hospitalization_id), "hospitalizations\n")
```

---

# Section 1 — Data Quality Checks

```{r data-quality}
# 1.1 Overall structure
cat("Rows:", nrow(df_raw), "\n")
cat("Columns:", ncol(df_raw), "\n")
cat("Unique patients:", n_distinct(df_raw$patient_id), "\n")
cat("Unique hospitalizations:", n_distinct(df_raw$hospitalization_id), "\n")
cat("Unique hospitals:", n_distinct(df_raw$hospital_id), "\n")

# 1.2 Duplicate rows — expect 0
dups <- df_raw %>%
  group_by(hospitalization_id, window_idx) %>%
  filter(n() > 1)
if (nrow(dups) > 0) {
  warning(nrow(dups), " duplicate rows detected. Review before proceeding.")
  print(dups)
} else {
  cat("Duplicate check: PASS\n")
}

# 1.3 Window gaps within hospitalizations
gap_check <- df_raw %>%
  group_by(hospitalization_id) %>%
  summarise(
    n_windows  = n(),
    min_window = min(window_idx),
    max_window = max(window_idx),
    has_gaps   = !identical(sort(window_idx), seq(min(window_idx), max(window_idx))),
    .groups    = "drop"
  ) %>%
  filter(has_gaps)

if (nrow(gap_check) > 0) {
  warning(nrow(gap_check), " hospitalizations have window gaps.")
  print(gap_check)
} else {
  cat("Window continuity check: PASS\n")
}

# 1.4 Outcome coding — expect only 0, 1, 2
cat("\nOutcome distribution (primary):\n")
print(df_raw %>% count(outcome_death_reintubate))

cat("\nOutcome distribution (secondary — NIPPV/CPAP):\n")
print(df_raw %>% count(outcome_death_reintubate_nippv_cpap))

# 1.5 Multiple events per hospitalization — expect 0 rows
multiple_events <- df_raw %>%
  group_by(hospitalization_id) %>%
  summarise(
    n_outcome1 = sum(outcome_death_reintubate == 1),
    n_outcome2 = sum(outcome_death_reintubate == 2),
    .groups    = "drop"
  ) %>%
  filter(n_outcome1 > 1 | n_outcome2 > 1)

if (nrow(multiple_events) > 0) {
  warning("Multiple events detected for ", nrow(multiple_events), " hospitalizations.")
  print(multiple_events)
} else {
  cat("Multiple events check: PASS\n")
}

# 1.6 Baseline completeness (window 0)
baseline_completeness <- df_raw %>%
  filter(window_idx == 0) %>%
  summarise(
    n              = n(),
    age_miss       = sum(is.na(age)),
    sex_miss       = sum(is.na(sex)),
    bmi_miss       = sum(is.na(bmi)),
    cci_miss       = sum(is.na(cci)),
    sofa_miss      = sum(is.na(sofa_extubation)),
    fio2_miss      = sum(is.na(hfno_fio2)),
    sf_miss        = sum(is.na(sf_ratio)),
    hfno_flow_miss = sum(is.na(hfno_flow_rate))
  )


cat("\nBaseline completeness:\n")
print(baseline_completeness)
write_csv(baseline_completeness, here("output_to_share/diagnostics/01_baseline_completeness.csv"))

# Flag BMI missingness early
if (baseline_completeness$bmi_miss > 0) {
  warning(baseline_completeness$bmi_miss,
          " hospitalizations have missing BMI — median imputation applied in weight models.")
}

```

---

# Section 2 — Cohort Description and Table 1

```{r cohort}
# 2.1 Final outcome per hospitalization
final_outcome_lookup <- df_raw %>%
  group_by(hospitalization_id) %>%
  arrange(window_idx) %>%
  summarise(
    has_failure   = any(outcome_death_reintubate == 1),
    has_discharge = any(outcome_death_reintubate == 2),
    max_window    = max(window_idx),
    n_windows     = n(),
    .groups       = "drop"
  )

# Check for conflicting outcome coding
if (any(final_outcome_lookup$has_failure & final_outcome_lookup$has_discharge)) {
  stop("ERROR: Some hospitalizations have both outcome=1 AND outcome=2. Fix before proceeding.")
}

final_outcome_lookup <- final_outcome_lookup %>%
  mutate(
    final_outcome = case_when(
      has_failure   ~ "Reintubation/Death",
      has_discharge ~ "Discharge Alive",
      TRUE          ~ "Censored"
    ),
    final_outcome = factor(final_outcome,
                           levels = c("Reintubation/Death", "Discharge Alive", "Censored"))
  )

# Outcome distribution table
outcome_summary <- final_outcome_lookup %>%
  count(final_outcome) %>%
  mutate(
    pct            = round(n / sum(n) * 100, 1),
    total_patients = sum(n),
    median_windows = median(final_outcome_lookup$n_windows),
    max_windows    = max(final_outcome_lookup$max_window)
  ) %>%
  select(
    `Final Outcome`              = final_outcome,
    N                            = n,
    `%`                          = pct,
    `Total Patients`             = total_patients,
    `Median Follow-up (windows)` = median_windows,
    `Max Follow-up (windows)`    = max_windows
  )

kable(outcome_summary, caption = "Cohort Outcome Distribution and Follow-up Summary")
write_csv(outcome_summary, here("output_to_share/tables/02_outcome_distribution.csv"))


#
# 2.2 Baseline dataset
df_baseline <- df_raw %>%
  filter(window_idx == 0) %>%
  select(hospitalization_id, patient_id,
         age, sex, race, bmi, cci, sofa_extubation,
         hfno_flow_rate, hfno_fio2, sf_ratio, rox_index, is_vaso) %>%
  rename(
    fio2_baseline     = hfno_fio2,
    sf_ratio_baseline = sf_ratio
  ) %>%
  # Median imputation for isolated baseline FiO2 missingness
  mutate(
    fio2_baseline = ifelse(is.na(fio2_baseline),
                           median(fio2_baseline, na.rm = TRUE),
                           fio2_baseline)
  ) %>%
  left_join(
    final_outcome_lookup %>% select(hospitalization_id, final_outcome, n_windows, max_window),
    by = "hospitalization_id"
  )

missing_w0 <- n_distinct(df_raw$hospitalization_id) - nrow(df_baseline)
if (missing_w0 > 0) {
  warning(missing_w0, " hospitalizations missing window 0 — excluded from baseline.")
}
# Missing/non-missing counts and flow rate descriptives per outcome
outcome_missingness <- final_outcome_lookup %>%
  left_join(df_baseline %>% select(hospitalization_id, hfno_flow_rate), 
            by = "hospitalization_id") %>%
  group_by(final_outcome) %>%
  summarise(
    n_total    = n(),
    n_missing  = sum(is.na(hfno_flow_rate)),
    n_nonmiss  = sum(!is.na(hfno_flow_rate)),
    mean_flow  = round(mean(hfno_flow_rate, na.rm = TRUE), 1),
    sd_flow    = round(sd(hfno_flow_rate,   na.rm = TRUE), 1),
    .groups    = "drop"
  )

write_csv(outcome_missingness, here("output_to_share/tables/02_outcome_missingness.csv"))
# 2.3 Table 1 by outcome
df_baseline %>%
  select(age, sex, bmi, cci, sofa_extubation,
         hfno_flow_rate, fio2_baseline, sf_ratio_baseline, rox_index,
         is_vaso, final_outcome) %>%
  tbl_summary(
    by = final_outcome,
    statistic = list(
      all_continuous()  ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = list(
      age               ~ "Age, years",
      sex               ~ "Sex",
      bmi               ~ "BMI, kg/m2",
      cci               ~ "Charlson Comorbidity Index",
      sofa_extubation   ~ "SOFA Score at Extubation",
      hfno_flow_rate    ~ "HFNO Flow Rate at Baseline, L/min",
      fio2_baseline     ~ "FiO2 at Baseline",
      sf_ratio_baseline ~ "S/F Ratio at Baseline",
      rox_index         ~ "ROX Index at Baseline",
      is_vaso           ~ "Vasopressors at Baseline"
    ),
    missing = "ifany"
  ) %>%
  add_overall(last = FALSE) %>%
  bold_labels() %>%
  as_tibble() %>%
  write_csv(here("output_to_share/tables/02_table1_by_outcome.csv"))

cat("Table 1 saved.\n")
```

---

# Section 3 — Exploratory Visualizations

```{r trajectories, fig.height=8}
df_plot <- df_raw %>%
  left_join(final_outcome_lookup %>% select(hospitalization_id, final_outcome),
            by = "hospitalization_id")

# 3.1 HFNO flow rate trajectory by outcome
df_flow_strat <- df_plot %>%
  filter(!is.na(hfno_flow_rate)) %>%
  group_by(window_idx, final_outcome) %>%
  summarise(
    n         = n(),
    mean_flow = mean(hfno_flow_rate),
    se        = sd(hfno_flow_rate) / sqrt(n),
    lci       = mean_flow - 1.96 * se,
    uci       = mean_flow + 1.96 * se,
    .groups   = "drop"
  )

write_csv(df_flow_strat, here("output_to_share/tables/03_hfno_trajectory_data.csv"))

p_flow <- ggplot(df_flow_strat,
                 aes(x = window_idx, y = mean_flow,
                     color = final_outcome, fill = final_outcome)) +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(aes(size = n), alpha = 0.8) +
  geom_vline(xintercept = 11.5, linetype = "dashed", color = "grey50") +
  scale_color_manual(values = outcome_colors, name = "Final Outcome") +
  scale_fill_manual(values  = outcome_colors, name = "Final Outcome") +
  scale_size_continuous(name = "N", range = c(1, 5)) +
  scale_x_continuous(breaks = seq(0, max(df_flow_strat$window_idx), by = 4),
                     labels = function(x) paste0(x * 4, "h")) +
  labs(title    = "HFNO Flow Rate Over Time by Final Outcome",
       subtitle = "Mean ± 95% CI | Dashed line = 48h",
       x = "Time Post-Extubation", y = "Mean HFNO Flow Rate (L/min)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank())

p_atrisk <- df_plot %>%
  group_by(window_idx, final_outcome) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = window_idx, y = n, fill = final_outcome)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.75) +
  scale_fill_manual(values = outcome_colors, guide = "none") +
  scale_x_continuous(breaks = seq(0, max(df_plot$window_idx), by = 4),
                     labels = function(x) paste0(x * 4, "h")) +
  labs(x = "Time Post-Extubation", y = "N at Risk") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank())

p_flow_combined <- p_flow / p_atrisk + plot_layout(heights = c(4, 1))
ggsave(here("output_to_share/figures/03_hfno_over_time.png"),
       plot = p_flow_combined, width = 11, height = 7.5, dpi = 300)
p_flow_combined

# 3.2 FiO2 trajectory
df_fio2_strat <- df_plot %>%
  filter(!is.na(hfno_fio2)) %>%
  group_by(window_idx, final_outcome) %>%
  summarise(
    n = n(), mean_fio2 = mean(hfno_fio2),
    se = sd(hfno_fio2) / sqrt(n),
    lci = mean_fio2 - 1.96 * se, uci = mean_fio2 + 1.96 * se,
    .groups = "drop"
  )

write_csv(df_fio2_strat, here("output_to_share/tables/03_fio2_trajectory_data.csv"))

p_fio2 <- ggplot(df_fio2_strat,
                 aes(x = window_idx, y = mean_fio2,
                     color = final_outcome, fill = final_outcome)) +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(aes(size = n), alpha = 0.8) +
  geom_vline(xintercept = 11.5, linetype = "dashed", color = "grey50") +
  scale_color_manual(values = outcome_colors, name = "Final Outcome") +
  scale_fill_manual(values  = outcome_colors, name = "Final Outcome") +
  scale_size_continuous(name = "N", range = c(1, 5)) +
  scale_x_continuous(breaks = seq(0, max(df_fio2_strat$window_idx), by = 4),
                     labels = function(x) paste0(x * 4, "h")) +
  labs(title = "FiO2 Over Time by Final Outcome", subtitle = "Mean ± 95% CI",
       x = "Time Post-Extubation", y = "Mean FiO2") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank())

ggsave(here("output_to_share/figures/03_fio2_over_time.png"), plot = p_fio2,
       width = 11, height = 6, dpi = 300)
p_fio2

# 3.3 SF ratio trajectory
df_sf_strat <- df_plot %>%
  filter(!is.na(sf_ratio)) %>%
  group_by(window_idx, final_outcome) %>%
  summarise(
    n = n(), mean_sf = mean(sf_ratio),
    se = sd(sf_ratio) / sqrt(n),
    lci = mean_sf - 1.96 * se, uci = mean_sf + 1.96 * se,
    .groups = "drop"
  )

write_csv(df_sf_strat, here("output_to_share/tables/03_sf_trajectory_data.csv"))

p_sf <- ggplot(df_sf_strat,
               aes(x = window_idx, y = mean_sf,
                   color = final_outcome, fill = final_outcome)) +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(aes(size = n), alpha = 0.8) +
  geom_vline(xintercept = 11.5, linetype = "dashed", color = "grey50") +
  scale_color_manual(values = outcome_colors, name = "Final Outcome") +
  scale_fill_manual(values  = outcome_colors, name = "Final Outcome") +
  scale_size_continuous(name = "N", range = c(1, 5)) +
  scale_x_continuous(breaks = seq(0, max(df_sf_strat$window_idx), by = 4),
                     labels = function(x) paste0(x * 4, "h")) +
  labs(title = "S/F Ratio Over Time by Final Outcome", subtitle = "Mean ± 95% CI",
       x = "Time Post-Extubation", y = "Mean S/F Ratio") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank())

ggsave(here("output_to_share/figures/03_sf_over_time.png"), plot = p_sf,
       width = 11, height = 6, dpi = 300)
p_sf

# 3.4 Baseline characteristics by outcome
p_baseline <- df_baseline %>%
  pivot_longer(cols = c(age, bmi, sofa_extubation, cci,
                        hfno_flow_rate, fio2_baseline, sf_ratio_baseline, is_vaso),
               names_to = "variable", values_to = "value") %>%
  mutate(var_label = case_when(
    variable == "age"               ~ "Age (years)",
    variable == "bmi"               ~ "BMI (kg/m2)",
    variable == "sofa_extubation"   ~ "SOFA at Extubation",
    variable == "cci"               ~ "Charlson Index",
    variable == "hfno_flow_rate"    ~ "HFNO Flow Rate (L/min)",
    variable == "fio2_baseline"     ~ "FiO2 at Baseline",
    variable == "sf_ratio_baseline" ~ "S/F Ratio at Baseline",
    variable == "is_vaso"           ~ "Vasopressors at Baseline"
  )) %>%
  ggplot(aes(x = final_outcome, y = value,
             fill = final_outcome, color = final_outcome)) +
  geom_boxplot(alpha = 0.4, outlier.shape = NA, width = 0.5) +
  geom_jitter(width = 0.15, size = 0.9, alpha = 0.5) +
  scale_fill_manual(values  = outcome_colors, guide = "none") +
  scale_color_manual(values = outcome_colors, guide = "none") +
  facet_wrap(~var_label, scales = "free_y", ncol = 3) +
  labs(title = "Baseline Characteristics by Final Outcome", x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(plot.title  = element_text(face = "bold"),
        strip.text  = element_text(face = "bold"),
        axis.text.x = element_text(angle = 25, hjust = 1))

ggsave(here("output_to_share/figures/03_baseline_characteristics.png"), plot = p_baseline,
       width = 12, height = 8, dpi = 300)
p_baseline

# 3.5 Overall HFNO trajectory
df_flow_overall <- df_raw %>%
  filter(!is.na(hfno_flow_rate)) %>%
  group_by(window_idx) %>%
  summarise(n = n(), mean_flow = mean(hfno_flow_rate),
            se = sd(hfno_flow_rate) / sqrt(n),
            lci = mean_flow - 1.96 * se, uci = mean_flow + 1.96 * se,
            .groups = "drop")

write_csv(df_flow_overall, here("output_to_share/tables/03_hfno_spline_data.csv"))

# Unstratified FiO2 trajectory (parallel to 03_hfno_spline_data)
df_fio2_overall <- df_raw %>%
  filter(!is.na(hfno_fio2)) %>%
  group_by(window_idx) %>%
  summarise(n = n(), mean_fio2 = mean(hfno_fio2),
            se  = sd(hfno_fio2) / sqrt(n),
            lci = mean_fio2 - 1.96 * se,
            uci = mean_fio2 + 1.96 * se,
            .groups = "drop")

write_csv(df_fio2_overall, here("output_to_share/tables/03_fio2_overall_data.csv"))

# Unstratified S/F trajectory
df_sf_overall <- df_raw %>%
  filter(!is.na(sf_ratio)) %>%
  group_by(window_idx) %>%
  summarise(n = n(), mean_sf = mean(sf_ratio),
            se  = sd(sf_ratio) / sqrt(n),
            lci = mean_sf - 1.96 * se,
            uci = mean_sf + 1.96 * se,
            .groups = "drop")

write_csv(df_sf_overall, here("output_to_share/tables/03_sf_overall_data.csv"))

```

---

# Section 4 — Exposure Cleaning and Person-Time Dataset

```{r person-time}
# LOCF applied BEFORE lag creation.
# Only window 0 is removed. Event rows preserved regardless of missingness.
# Cleaning protocol:
#   Step 1: Observed value -> use as-is
#   Step 2: Isolated gap -> LOCF
#   Step 3: Sustained missing, non-terminal -> 0 (weaned off HFNO)
#   Step 4: Missing at event row -> LOCF from last observed

df_clean <- df_raw %>%
  group_by(hospitalization_id) %>%
  arrange(window_idx) %>%
  mutate(
    is_terminal          = outcome_death_reintubate %in% c(1, 2),
    hfno_flow_rate_locf  = na.locf(hfno_flow_rate, na.rm = FALSE),
    hfno_flow_rate_clean = case_when(
      !is.na(hfno_flow_rate)      ~ hfno_flow_rate,
      is_terminal                  ~ hfno_flow_rate_locf,
      !is.na(hfno_flow_rate_locf) ~ hfno_flow_rate_locf,
      TRUE                         ~ 0
    ),
    hfno_fio2_clean = case_when(
      !is.na(hfno_fio2) ~ hfno_fio2,
      TRUE               ~ na.locf(hfno_fio2, na.rm = FALSE)
    ),
    sf_ratio_clean = case_when(
      !is.na(sf_ratio) ~ sf_ratio,
      TRUE              ~ na.locf(sf_ratio, na.rm = FALSE)
    ),
    high_dose_clean  = as.integer(hfno_flow_rate_clean >= 40),
    window_idx_num   = as.numeric(window_idx),
    sex              = factor(sex),
    race             = factor(race),
    event_cs1        = if_else(outcome_death_reintubate == 1, 1L, 0L),
    event_cs2        = if_else(outcome_death_reintubate == 2, 1L, 0L)
  ) %>%
  ungroup()

# 4.2 Person-time dataset with lags (created after cleaning)
df_msm <- df_clean %>%
  group_by(hospitalization_id) %>%
  arrange(window_idx) %>%
  mutate(
    hfno_flow_rate_lag1 = lag(hfno_flow_rate_clean),
    high_dose_lag1      = lag(high_dose_clean),
    fio2_lag1           = lag(hfno_fio2_clean),
    sf_ratio_lag1       = lag(sf_ratio_clean),
    rox_index_lag1      = lag(rox_index),
    is_event_row        = outcome_death_reintubate %in% c(1, 2)
  ) %>%
  ungroup() %>%
  filter(window_idx > 0) %>%
  left_join(
    final_outcome_lookup %>% select(hospitalization_id, final_outcome),
    by = "hospitalization_id"
  )

# 4.3 Event preservation check
window0_events <- df_clean %>%
  filter(window_idx == 0 & outcome_death_reintubate %in% c(1, 2)) %>%
  select(hospitalization_id, outcome_death_reintubate)

if (nrow(window0_events) > 0) {
  message(nrow(window0_events), " patient(s) had event at window 0 — excluded by design. ",
          "Verify event timestamps in source data.")
  print(window0_events)
}

last_row_outcomes <- df_msm %>%
  group_by(hospitalization_id) %>%
  slice_tail(n = 1) %>%
  ungroup() %>%
  count(outcome_death_reintubate)

expected_failures  <- sum(final_outcome_lookup$has_failure)
expected_discharge <- sum(final_outcome_lookup$has_discharge)
observed_failures  <- last_row_outcomes %>% filter(outcome_death_reintubate == 1) %>% pull(n)
observed_discharge <- last_row_outcomes %>% filter(outcome_death_reintubate == 2) %>% pull(n)
if (length(observed_failures)  == 0) observed_failures  <- 0
if (length(observed_discharge) == 0) observed_discharge <- 0

n_window0_excluded <- nrow(window0_events)

if (abs(observed_failures  - expected_failures)  > n_window0_excluded |
    abs(observed_discharge - expected_discharge) > n_window0_excluded) {
  warning("Event counts do not match after person-time expansion. Investigate before proceeding.")
} else {
  message("Event preservation check: PASS (", n_window0_excluded,
          " window-0 event(s) excluded by design)")
}

cat("Person-time rows:", nrow(df_msm), "\n")
cat("Unique hospitalizations:", n_distinct(df_msm$hospitalization_id), "\n")
```

---

# Section 5 — Outcome Validation

```{r outcome-validation, fig.height=5}
events_by_window <- df_msm %>%
  group_by(window_idx) %>%
  summarise(
    n_at_risk = n(),
    n_event1  = sum(event_cs1),
    n_event2  = sum(event_cs2),
    n_ongoing = sum(outcome_death_reintubate == 0),
    .groups   = "drop"
  )

write_csv(events_by_window, here("output_to_share/tables/05_events_by_window.csv"))

p_events <- df_msm %>%
  filter(outcome_death_reintubate %in% c(1, 2)) %>%
  mutate(outcome_label = factor(
    ifelse(outcome_death_reintubate == 1, "Failure (Reintubation/Death)", "Discharge Alive"),
    levels = c("Failure (Reintubation/Death)", "Discharge Alive")
  )) %>%
  count(window_idx, outcome_label) %>%
  mutate(hour = window_idx * 4) %>%
  ggplot(aes(x = hour, y = n, fill = outcome_label)) +
  geom_col(position = "dodge", alpha = 0.85, width = 3) +
  geom_vline(xintercept = 48, linetype = "dashed", color = "grey50") +
  scale_fill_manual(
    values = c("Failure (Reintubation/Death)" = "#C0392B",
               "Discharge Alive"              = "#27AE60"),
    name = NULL
  ) +
  scale_x_continuous(breaks = seq(0, max(df_msm$window_idx) * 4, by = 16),
                     labels = function(x) paste0(x, "h")) +
  scale_y_continuous(breaks = breaks_pretty()) +
  labs(title = "Events by Time Window",
       subtitle = "Dashed line = 48h post-extubation",
       x = "Time Post-Extubation", y = "Number of Events") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position = "bottom")

ggsave(here("output_to_share/figures/05_events_by_window.png"),
       plot = p_events, width = 10, height = 5, dpi = 300)
p_events
```

---

# Section 6 — Baseline Balance Diagnostics

```{r baseline-diagnostics, fig.height=5}
# 6.1 Baseline HFNO flow distribution by outcome
p_baseline_dist <- df_baseline %>%
  filter(!is.na(hfno_flow_rate)) %>%
  ggplot(aes(x = hfno_flow_rate, fill = final_outcome)) +
  geom_histogram(binwidth = 5, position = "stack", alpha = 0.85, color = "white") +
  scale_fill_manual(values = outcome_colors, name = "Final Outcome") +
  labs(title   = "Distribution of Baseline HFNO Flow Rate by Outcome",
       x = "HFNO Flow Rate (L/min)", y = "Count") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

ggsave(here("output_to_share/figures/06_baseline_hfno_distribution.png"),
       plot = p_baseline_dist, width = 8, height = 5, dpi = 300)
p_baseline_dist

# 6.2 Correlation matrix — confounders vs exposure
# Justifies MSM and pre-specified confounder selection
cor_table <- df_baseline %>%
  select(hfno_flow_rate, sofa_extubation, age, bmi, cci,
         fio2_baseline, sf_ratio_baseline, rox_index, is_vaso) %>%
  mutate(is_vaso = as.numeric(is_vaso)) %>%
  cor(use = "pairwise.complete.obs") %>%
  round(2)

write_csv(
  as_tibble(cor_table, rownames = "variable"),
  here("output_to_share/tables/06_correlation_matrix.csv"
))

cat("Correlation of confounders with HFNO flow rate (justifies MSM):\n")
print(cor_table["hfno_flow_rate", , drop = FALSE])

# 6.3 Baseline scatter plots

cor_hfno_fio2 <- cor.test(df_baseline$hfno_flow_rate, df_baseline$fio2_baseline)
cor_hfno_sf   <- cor.test(df_baseline$hfno_flow_rate, df_baseline$sf_ratio_baseline)

p_hfno_fio2 <- df_baseline %>%
  filter(!is.na(hfno_flow_rate), !is.na(fio2_baseline)) %>%
  ggplot(aes(x = fio2_baseline, y = hfno_flow_rate, color = final_outcome)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(aes(group = 1), method = "lm", se = TRUE,
              color = "black", linetype = "dashed", linewidth = 0.8) +
  scale_color_manual(values = outcome_colors, name = "Final Outcome") +
  labs(title    = "Baseline HFNO Flow Rate vs FiO2",
       subtitle = sprintf("Pearson r = %.2f, p = %.4f",
                          cor_hfno_fio2$estimate, cor_hfno_fio2$p.value),
       x = "FiO2 at Baseline", y = "HFNO Flow Rate (L/min)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")

p_hfno_sf <- df_baseline %>%
  filter(!is.na(hfno_flow_rate), !is.na(sf_ratio_baseline)) %>%
  ggplot(aes(x = sf_ratio_baseline, y = hfno_flow_rate, color = final_outcome)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(aes(group = 1), method = "lm", se = TRUE,
              color = "black", linetype = "dashed", linewidth = 0.8) +
  scale_color_manual(values = outcome_colors, name = "Final Outcome") +
  labs(title    = "Baseline HFNO Flow Rate vs S/F Ratio",
       subtitle = sprintf("Pearson r = %.2f, p = %.4f",
                          cor_hfno_sf$estimate, cor_hfno_sf$p.value),
       x = "S/F Ratio at Baseline", y = "HFNO Flow Rate (L/min)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")

ggsave(here("output_to_share/figures/06_baseline_hfno_vs_fio2.png"), plot = p_hfno_fio2,
       width = 9, height = 6, dpi = 300)
ggsave(here("output_to_share/figures/06_baseline_hfno_vs_sf.png"),   plot = p_hfno_sf,
       width = 9, height = 6, dpi = 300)

p_hfno_fio2 + p_hfno_sf
```

---

# Section 7 — MSM Weights

**Pre-specified weight model (from pilot site elimination analysis):**

- **Denominator:** flow_lag1 + time spline + sofa_extubation + age + bmi + FiO2_lag1
- **Numerator:** flow_lag1 + time spline + sofa_extubation + age + bmi
- **TV confounder:** FiO2_lag1 only (SF ratio and ROX index collinear; tested in sensitivity)
- Sex and CCI dropped from both models (r < 0.15 with flow rate; OR unchanged on elimination)

```{r msm-weights}
# Join baseline covariates to person-time data
df_msm <- df_msm %>%
  left_join(
    df_baseline %>% select(hospitalization_id, fio2_baseline, sf_ratio_baseline),
    by = "hospitalization_id"
  )

# Impute residual lag missingness with column median (prevents row loss in lm)
df_weights <- df_msm %>%
  mutate(
    fio2_lag1_imp = ifelse(is.na(fio2_lag1),
                           median(fio2_lag1, na.rm = TRUE), fio2_lag1),
    sf_lag1_imp   = ifelse(is.na(sf_ratio_lag1),
                           median(sf_ratio_lag1, na.rm = TRUE), sf_ratio_lag1),
    flow_lag1_imp = ifelse(is.na(hfno_flow_rate_lag1),
                           median(hfno_flow_rate_lag1, na.rm = TRUE),
                           hfno_flow_rate_lag1)
  )
#Impute BMI if needed - same process for missing above 
df_weights <- df_msm %>%
  mutate(
    # existing imputations
    fio2_lag1_imp = ifelse(is.na(fio2_lag1),
                           median(fio2_lag1, na.rm = TRUE), fio2_lag1),
    sf_lag1_imp   = ifelse(is.na(sf_ratio_lag1),
                           median(sf_ratio_lag1, na.rm = TRUE), sf_ratio_lag1),
    flow_lag1_imp = ifelse(is.na(hfno_flow_rate_lag1),
                           median(hfno_flow_rate_lag1, na.rm = TRUE),
                           hfno_flow_rate_lag1),
    # new — BMI imputation
    bmi_imp       = ifelse(is.na(bmi),
                           median(bmi, na.rm = TRUE), bmi)
  )
# 7.2 Denominator model
denom_model <- lm(
  hfno_flow_rate_clean ~
    flow_lag1_imp +
    ns(window_idx_num, df = 3) +
    sofa_extubation + age + bmi_imp +
    fio2_lag1_imp,
  data = df_weights
)

# 7.3 Numerator model
numer_model <- lm(
  hfno_flow_rate_clean ~
    flow_lag1_imp +
    ns(window_idx_num, df = 3) +
    sofa_extubation + age + bmi_imp,
  data = df_weights
)

cat("Denominator model R²:", round(summary(denom_model)$r.squared, 3), "\n")
cat("Numerator model R²:  ", round(summary(numer_model)$r.squared, 3), "\n")

# 7.4 Per-interval stabilized weights
denom_sd <- summary(denom_model)$sigma
numer_sd <- summary(numer_model)$sigma

df_weights <- df_weights %>%
  mutate(
    pred_denom = predict(denom_model, newdata = df_weights),
    pred_numer = predict(numer_model, newdata = df_weights),
    dens_denom = dnorm(hfno_flow_rate_clean, mean = pred_denom, sd = denom_sd),
    dens_numer = dnorm(hfno_flow_rate_clean, mean = pred_numer, sd = numer_sd),
    w_interval = ifelse(dens_denom > 1e-10, dens_numer / dens_denom, 1)
  )

# 7.5 Cumulative weights
df_weights <- df_weights %>%
  group_by(hospitalization_id) %>%
  arrange(window_idx_num) %>%
  mutate(w_msm_raw = cumprod(w_interval)) %>%
  ungroup()

# 7.6 Pre-truncation diagnostics
weight_diag_pre <- df_weights %>%
  summarise(
    mean_w    = round(mean(w_msm_raw,     na.rm = TRUE), 3),
    median_w  = round(median(w_msm_raw,   na.rm = TRUE), 3),
    sd_w      = round(sd(w_msm_raw,       na.rm = TRUE), 3),
    min_w     = round(min(w_msm_raw,      na.rm = TRUE), 3),
    max_w     = round(max(w_msm_raw,      na.rm = TRUE), 3),
    p99_w     = round(quantile(w_msm_raw, 0.99, na.rm = TRUE), 3),
    n_extreme = sum(w_msm_raw > 10, na.rm = TRUE)
  )

cat("\nWeight diagnostics (pre-truncation):\n")
print(weight_diag_pre)

if (weight_diag_pre$max_w > 50) {
  warning("Max weight >50 — flag for coordinating site before proceeding.")
}

# 7.7 Truncate at 1st/99th percentile
w_p01 <- quantile(df_weights$w_msm_raw, .01, na.rm = TRUE)
w_p99 <- quantile(df_weights$w_msm_raw, .99, na.rm = TRUE)

df_weights <- df_weights %>%
  mutate(w_msm_trunc = pmin(pmax(w_msm_raw, w_p01), w_p99))

weight_diag_post <- df_weights %>%
  summarise(
    mean_w   = round(mean(w_msm_trunc,   na.rm = TRUE), 3),
    median_w = round(median(w_msm_trunc, na.rm = TRUE), 3),
    sd_w     = round(sd(w_msm_trunc,     na.rm = TRUE), 3),
    max_w    = round(max(w_msm_trunc,    na.rm = TRUE), 2)
  )

cat("\nWeight diagnostics (post-truncation):\n")
print(weight_diag_post)

# 7.8 ESS by time window
ess_by_window <- df_weights %>%
  group_by(window_idx_num) %>%
  summarise(
    n        = n(),
    mean_w   = round(mean(w_msm_trunc,   na.rm = TRUE), 3),
    median_w = round(median(w_msm_trunc, na.rm = TRUE), 3),
    ess      = sum(w_msm_trunc, na.rm = TRUE)^2 / sum(w_msm_trunc^2, na.rm = TRUE),
    ess_pct  = round(ess / n * 100, 1),
    max_w    = round(max(w_msm_trunc,    na.rm = TRUE), 2),
    .groups  = "drop"
  )

write_csv(ess_by_window, here("output_to_share/models/07_weight_diagnostics.csv"))
```

```{r weight-plot, fig.height=5}
p_weights <- ggplot(df_weights, aes(x = w_msm_raw)) +
  geom_histogram(bins = 50, fill = "#2E75B6", alpha = 0.7) +
  geom_vline(xintercept = c(w_p01, w_p99),
             linetype = "dashed", color = "#C0392B") +
  scale_x_log10() +
  labs(title    = "Cumulative IPT Weight Distribution",
       subtitle = "Dashed lines = 1st/99th percentile truncation points",
       x = "Weight (log scale)", y = "Count") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))

ggsave(here("output_to_share/figures/07_weight_distribution.png"),
       plot = p_weights, width = 8, height = 5, dpi = 300)
p_weights

weight_distribution_export <- df_weights %>%
  select(hospitalization_id, window_idx_num, w_msm_raw, w_msm_trunc) %>%
  summarise(
    n          = n(),
    mean_raw   = round(mean(w_msm_raw,   na.rm = TRUE), 4),
    sd_raw     = round(sd(w_msm_raw,     na.rm = TRUE), 4),
    p01_raw    = round(quantile(w_msm_raw, .01, na.rm = TRUE), 4),
    p99_raw    = round(quantile(w_msm_raw, .99, na.rm = TRUE), 4),
    max_raw    = round(max(w_msm_raw,     na.rm = TRUE), 4),
    mean_trunc = round(mean(w_msm_trunc, na.rm = TRUE), 4),
    sd_trunc   = round(sd(w_msm_trunc,   na.rm = TRUE), 4),
    max_trunc  = round(max(w_msm_trunc,  na.rm = TRUE), 4)
  )

write_csv(weight_distribution_export, here("output_to_share/models/07_weight_distribution.csv"))

df_msm_final <- df_weights
```

---
# Section 8 — Primary Outcome Model

**Model specification:** Outcome model includes exposure + time spline only. All confounding adjustment (sofa_extubation, age, bmi, time-varying FiO2) is handled by the MSM weights from Section 7. Parsimonious outcome model justified by pilot site elimination analysis and power constraints.

```{r primary-model}
# Convert outcomes to factors — multinom's Hessian computation (multinomHess)
# fails on numeric vectors due to internal dimension mismatch in probs matrix.
# Applied here so factor state is consistent across df_msm_final, df_weights, and df_weights_sf
df_msm_final <- df_msm_final %>%
  mutate(
    outcome_death_reintubate            = factor(outcome_death_reintubate,            levels = c(0, 1, 2)),
    outcome_death_reintubate_nippv_cpap = factor(outcome_death_reintubate_nippv_cpap, levels = c(0, 1, 2))
  )

df_weights <- df_weights %>%
  mutate(
    outcome_death_reintubate            = factor(outcome_death_reintubate,            levels = c(0, 1, 2)),
    outcome_death_reintubate_nippv_cpap = factor(outcome_death_reintubate_nippv_cpap, levels = c(0, 1, 2))
  )

# 8.1 Unweighted model (conventional — comparator)
df_model <- df_msm_final %>% filter(!is.na(hfno_flow_rate_lag1))

cat("Rows entering outcome model:", nrow(df_model), "\n")
cat("Non-missing weights:", sum(!is.na(df_model$w_msm_trunc)), "\n")

model_unweighted <- nnet::multinom(
  outcome_death_reintubate ~
    hfno_flow_rate_lag1 +
    ns(window_idx_num, df = 3),
  data  = df_model,
  trace = FALSE
)

coef_unweighted <- summary(model_unweighted)$coefficients
se_unweighted   <- summary(model_unweighted)$standard.errors

# Confirm row names after factoring — should be "1" and "2"
cat("Coefficient row names:", rownames(coef_unweighted), "\n")

# 8.2 Weighted model (MSM causal estimate)
model_weighted <- nnet::multinom(
  outcome_death_reintubate ~
    hfno_flow_rate_lag1 +
    ns(window_idx_num, df = 3),
  data    = df_model,
  weights = df_model$w_msm_trunc,
  trace   = FALSE
)

coef_weighted <- summary(model_weighted)$coefficients
se_weighted   <- summary(model_weighted)$standard.errors

# 8.3 Full coefficient tables
full_coef_unweighted <- as_tibble(coef_unweighted, rownames = "outcome") %>%
  pivot_longer(-outcome, names_to = "term", values_to = "coef") %>%
  left_join(
    as_tibble(se_unweighted, rownames = "outcome") %>%
      pivot_longer(-outcome, names_to = "term", values_to = "se"),
    by = c("outcome", "term")
  ) %>%
  mutate(model    = "unweighted",
         or       = exp(coef),
         or_lower = exp(coef - 1.96 * se),
         or_upper = exp(coef + 1.96 * se))

full_coef_weighted <- as_tibble(coef_weighted, rownames = "outcome") %>%
  pivot_longer(-outcome, names_to = "term", values_to = "coef") %>%
  left_join(
    as_tibble(se_weighted, rownames = "outcome") %>%
      pivot_longer(-outcome, names_to = "term", values_to = "se"),
    by = c("outcome", "term")
  ) %>%
  mutate(model    = "weighted_msm",
         or       = exp(coef),
         or_lower = exp(coef - 1.96 * se),
         or_upper = exp(coef + 1.96 * se))

# 8.4 Primary results table (key deliverable)
results_comparison <- tibble(
  outcome = c("Failure (Reintubation/Death)", "Failure (Reintubation/Death)",
              "Discharge Alive",              "Discharge Alive"),
  model   = c("Unweighted", "Weighted (MSM)",
              "Unweighted", "Weighted (MSM)"),
  coef_per_lpm = c(
    coef_unweighted["1", "hfno_flow_rate_lag1"],
    coef_weighted["1",   "hfno_flow_rate_lag1"],
    coef_unweighted["2", "hfno_flow_rate_lag1"],
    coef_weighted["2",   "hfno_flow_rate_lag1"]
  ),
  or_per_10lpm = exp(coef_per_lpm * 10),
  se_coef = c(
    se_unweighted["1", "hfno_flow_rate_lag1"],
    se_weighted["1",   "hfno_flow_rate_lag1"],
    se_unweighted["2", "hfno_flow_rate_lag1"],
    se_weighted["2",   "hfno_flow_rate_lag1"]
  ),
  or_lower = exp((coef_per_lpm - 1.96 * se_coef) * 10),
  or_upper = exp((coef_per_lpm + 1.96 * se_coef) * 10)
) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))

kable(results_comparison, caption = "Primary Results: OR per 10 L/min Increase in HFNO Flow Rate")
write_csv(results_comparison,   here("output_to_share/models/08_model3_exposure_effect.csv"))
write_csv(full_coef_unweighted, here("output_to_share/models/08_model3_unweighted_full.csv"))
write_csv(full_coef_weighted,   here("output_to_share/models/08_model3_weighted_full.csv"))
```

---

# Section 9 — Clustering Adjustment

```{r clustering}
# Safe default — overwritten by each branch below
clustering_method <- "Unknown — check clustering detection"

# 9.1 Detect clustering structure
# Guard: create placeholder variables if hospital/ICU identifiers are absent
has_hospital_var <- "hospital_id"         %in% names(df_msm_final)
has_icu_var      <- "extubation_icu_type" %in% names(df_msm_final)

if (!has_hospital_var) {
  warning("hospital_id not found — assuming single hospital.")
  df_msm_final$hospital_id <- "single_hospital"
}
if (!has_icu_var) {
  warning("extubation_icu_type not found — assuming single ICU type.")
  df_msm_final$extubation_icu_type <- "single_icu"
}

n_hospitals <- n_distinct(df_msm_final$hospital_id[!is.na(df_msm_final$hospital_id)])
n_icus      <- n_distinct(df_msm_final$extubation_icu_type[!is.na(df_msm_final$extubation_icu_type)])

n_na_cluster <- sum(is.na(df_msm_final$hospital_id) | is.na(df_msm_final$extubation_icu_type))
if (n_na_cluster > 0) {
  warning(n_na_cluster, " person-time rows have missing hospital_id or extubation_icu_type — excluded from clustering model.")
}

cat("Clustering structure at this site:\n")
cat("  Hospitals:", n_hospitals, "\n")
cat("  ICU types:", n_icus,      "\n")

# Create aligned clustering dataset — filtered and factored consistently
df_cluster <- df_msm_final %>%
  filter(!is.na(hfno_flow_rate_lag1),
         !is.na(hospital_id),
         !is.na(extubation_icu_type)) %>%
  mutate(outcome_death_reintubate = factor(outcome_death_reintubate, levels = c(0, 1, 2)))

cat("Rows entering clustering model:", nrow(df_cluster), "\n")

# 9.2 Fit clustering-adjusted model
if (n_hospitals == 1 & n_icus == 1) {

  cat("Single hospital, single ICU type — no clustering adjustment needed.\n")
  model_final       <- model_weighted
  clustering_method <- "None — single hospital, single ICU type"

} else if (n_hospitals == 1 & n_icus > 1) {

  cat("Single hospital, multiple ICU types — applying ICU-level random intercept.\n")
  model_final <- tryCatch({
    mclogit::mblogit(
      outcome_death_reintubate ~
        hfno_flow_rate_lag1 +
        ns(window_idx_num, df = 3),
      random  = ~ 1 | extubation_icu_type,
      data    = df_cluster,
      weights = df_cluster$w_msm_trunc
    )
  }, error = function(e) {
    cat("mblogit failed:", conditionMessage(e), "\n")
    cat("Falling back to primary weighted model.\n")
    model_weighted
  })
  clustering_method <- paste0("ICU-level random intercept (n ICU types=", n_icus, ")")

} else if (n_hospitals > 1) {

  cat("Multiple hospitals — attempting nested ICU-within-hospital random intercept.\n")
  model_final <- tryCatch({
    mclogit::mblogit(
      outcome_death_reintubate ~
        hfno_flow_rate_lag1 +
        ns(window_idx_num, df = 3),
      random  = ~ 1 | hospital_id / extubation_icu_type,
      data    = df_cluster,
      weights = df_cluster$w_msm_trunc
    )
  }, error = function(e) {
    cat("Nested mblogit failed:", conditionMessage(e), "\n")
    cat("Attempting hospital-only random intercept as fallback.\n")
    tryCatch({
      mclogit::mblogit(
        outcome_death_reintubate ~
          hfno_flow_rate_lag1 +
          ns(window_idx_num, df = 3),
        random  = ~ 1 | hospital_id,
        data    = df_cluster,
        weights = df_cluster$w_msm_trunc
      )
    }, error = function(e2) {
      cat("Hospital-only mblogit also failed:", conditionMessage(e2), "\n")
      cat("Falling back to primary weighted model.\n")
      model_weighted
    })
  })
  clustering_method <- paste0("Nested ICU-within-hospital random intercept (n hospitals=",
                               n_hospitals, ", n ICU types=", n_icus, ") — or fallback")
}

if (!exists("model_final")) {
  model_final       <- model_weighted
  clustering_method <- "Fallback — model_final not created; using primary weighted model"
}

# 9.3 Extract clustering-adjusted results
if (n_hospitals == 1 & n_icus == 1) {

  results_final <- results_comparison %>%
    mutate(clustering = clustering_method)

} else {

  coef_final    <- coef(model_final)
  se_final      <- sqrt(diag(vcov(model_final)))
  exposure_rows <- grepl("hfno_flow_rate_lag1", names(coef_final))

  results_final <- tibble(
    term         = names(coef_final)[exposure_rows],
    coef         = coef_final[exposure_rows],
    se           = se_final[exposure_rows],
    or_per_10lpm = round(exp(coef * 10), 3),
    or_lower     = round(exp((coef - 1.96 * se) * 10), 3),
    or_upper     = round(exp((coef + 1.96 * se) * 10), 3),
    clustering   = clustering_method
  )
}

kable(results_final, caption = "Clustering-Adjusted Results")
write_csv(results_final, here("output_to_share/models/09_clustering_adjusted_results.csv"))

# 9.4 ICC (only when random effects model was fitted)
if (n_hospitals > 1 | n_icus > 1) {
  tryCatch({

    if (inherits(model_final, "mblogit") && !is.null(model_final$VarCov)) {

      vc_list <- model_final$VarCov
      icc_rows <- lapply(names(vc_list), function(level) {
        var_u <- as.numeric(vc_list[[level]][1, 1])
        icc   <- var_u / (var_u + (pi^2 / 3))
        tibble(
          clustering_level  = level,
          var_intercept     = round(var_u, 4),
          icc               = round(icc,   4),
          clustering_method = clustering_method
        )
      })

      icc_out <- bind_rows(icc_rows)
      print(icc_out)
      cat("ICC > 0.05 indicates meaningful clustering warranting random effects.\n")
      write_csv(icc_out, here("output_to_share/models/09_icc.csv"))

    } else {
      cat("model_final is not an mblogit object (fallback to multinom fired) — ICC not applicable.\n")
      write_csv(
        tibble(clustering_level  = "none",
               var_intercept     = NA,
               icc               = NA,
               clustering_method = clustering_method),
        here("output_to_share/models/09_icc.csv")
      )
    }

  }, error = function(e) {
    cat("ICC extraction failed:", conditionMessage(e), "\n")
    cat("Run str(model_final$VarCov) manually to inspect variance components.\n")
  })
}
```

---

# Section 10 — Sensitivity Analyses

```{r sensitivity-secondary}
# 10.1 Secondary outcome: NIPPV/CPAP included in failure definition
# NOTE: Spline df may need adjustment if events are sparse in early windows.
# Check the event count table below before interpreting model results.

n_secondary_events <- table(df_msm_final$outcome_death_reintubate_nippv_cpap)
cat("Secondary outcome event distribution:\n")
print(n_secondary_events)

model_secondary <- nnet::multinom(
  outcome_death_reintubate_nippv_cpap ~
    hfno_flow_rate_lag1 +
    ns(window_idx_num, df = 2),
  data    = df_model,
  weights = df_model$w_msm_trunc,
  trace   = FALSE
)

secondary_results <- tidy(model_secondary) %>%
  filter(term == "hfno_flow_rate_lag1") %>%
  mutate(
    or_per_10lpm = round(exp(estimate * 10), 3),
    or_lower     = round(exp((estimate - 1.96 * std.error) * 10), 3),
    or_upper     = round(exp((estimate + 1.96 * std.error) * 10), 3),
    p_value      = round(p.value, 4),
    sensitivity  = "Secondary outcome (NIPPV/CPAP as failure)"
  ) %>%
  select(sensitivity, y.level, or_per_10lpm, or_lower, or_upper, p_value)

kable(secondary_results, caption = "10.1 Secondary Outcome Sensitivity")
write_csv(secondary_results, here("output_to_share/models/10_sensitivity_secondary_outcome.csv"))
```

```{r sensitivity-untruncated}
# 10.2 Untruncated weights
model_untruncated <- nnet::multinom(
  outcome_death_reintubate ~
    hfno_flow_rate_lag1 +
    ns(window_idx_num, df = 3),
  data    = df_model,
  weights = df_model$w_msm_raw,
  trace   = FALSE
)

untrunc_results <- tidy(model_untruncated) %>%
  filter(term == "hfno_flow_rate_lag1") %>%
  mutate(
    or_per_10lpm = round(exp(estimate * 10), 3),
    or_lower     = round(exp((estimate - 1.96 * std.error) * 10), 3),
    or_upper     = round(exp((estimate + 1.96 * std.error) * 10), 3),
    p_value      = round(p.value, 4),
    sensitivity  = "Untruncated weights"
  ) %>%
  select(sensitivity, y.level, or_per_10lpm, or_lower, or_upper, p_value)

kable(untrunc_results, caption = "10.2 Untruncated Weights Sensitivity")
write_csv(untrunc_results, here("output_to_share/models/10_sensitivity_untruncated_weights.csv"))
```

```{r sensitivity-sf}
# 10.3 Sensitivity A: SF ratio as TV confounder instead of FiO2
denom_model_sf <- lm(
  hfno_flow_rate_clean ~
    flow_lag1_imp +
    ns(window_idx_num, df = 3) +
    sofa_extubation + age + bmi_imp +
    sf_lag1_imp,
  data = df_weights
)

denom_sd_sf <- summary(denom_model_sf)$sigma

# df_weights already has outcome factored from Section 8
df_weights_sf <- df_weights %>%
  mutate(
    pred_denom_sf = predict(denom_model_sf, newdata = df_weights),
    dens_denom_sf = dnorm(hfno_flow_rate_clean, mean = pred_denom_sf, sd = denom_sd_sf),
    # dens_numer reused from primary model — numerator is invariant across denominator sensitivities by design
    w_interval_sf = ifelse(dens_denom_sf > 1e-10, dens_numer / dens_denom_sf, 1)
  ) %>%
  group_by(hospitalization_id) %>%
  arrange(window_idx_num) %>%
  mutate(w_msm_sf_raw = cumprod(w_interval_sf)) %>%
  ungroup()

# Global quantile truncation (must be outside group_by)
sf_p01 <- quantile(df_weights_sf$w_msm_sf_raw, .01, na.rm = TRUE)
sf_p99 <- quantile(df_weights_sf$w_msm_sf_raw, .99, na.rm = TRUE)

df_weights_sf <- df_weights_sf %>%
  mutate(w_msm_sf_trunc = pmin(pmax(w_msm_sf_raw, sf_p01), sf_p99))

# Aligned dataset for SF sensitivity model
df_model_sf <- df_weights_sf %>% filter(!is.na(hfno_flow_rate_lag1))

model_sf_confounder <- nnet::multinom(
  outcome_death_reintubate ~
    hfno_flow_rate_lag1 +
    ns(window_idx_num, df = 3),
  data    = df_model_sf,
  weights = df_model_sf$w_msm_sf_trunc,
  trace   = FALSE
)

sf_results <- tidy(model_sf_confounder) %>%
  filter(term == "hfno_flow_rate_lag1") %>%
  mutate(
    or_per_10lpm = round(exp(estimate * 10), 3),
    or_lower     = round(exp((estimate - 1.96 * std.error) * 10), 3),
    or_upper     = round(exp((estimate + 1.96 * std.error) * 10), 3),
    p_value      = round(p.value, 4),
    sensitivity  = "SF ratio as TV confounder"
  ) %>%
  select(sensitivity, y.level, or_per_10lpm, or_lower, or_upper, p_value)

kable(sf_results, caption = "10.3 SF Ratio as TV Confounder Sensitivity")
write_csv(sf_results, here("output_to_share/models/10_sensitivity_sf_confounder.csv"))
```

```{r sensitivity-spline}
# 10.4 Nonlinearity test: spline on continuous exposure
model_spline <- nnet::multinom(
  outcome_death_reintubate ~
    ns(hfno_flow_rate_lag1, df = 3) +
    ns(window_idx_num, df = 3),
  data    = df_model,
  weights = df_model$w_msm_trunc,
  trace   = FALSE
)

spline_results <- tidy(model_spline) %>%
  filter(grepl("hfno_flow_rate_lag1", term)) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)),
         sensitivity = "Spline on continuous exposure")

kable(spline_results, caption = "10.4 Nonlinearity Test (Spline on Exposure)")
write_csv(spline_results, here("output_to_share/models/10_sensitivity_spline_exposure.csv"))
```

---

# Session Info

```{r session-info}
writeLines(capture.output(sessionInfo()), here("output_to_share/session_info.txt"))
sessionInfo()
```

```

---

# Output File Summary

The following files have been saved to the `output_to_share/` directory for return to the coordinating site:

**Tables (return all):**

- `output_to_share/tables/02_outcome_distribution.csv` — cohort outcome counts
- `output_to_share/tables/02_table1_by_outcome.csv` — Table 1
- `output_to_share/tables/02_outcome_missingness.csv`
- `output_to_share/tables/03_hfno_trajectory_data.csv` — flow rate by window
- `output_to_share/tables/03_fio2_trajectory_data.csv` — FiO2 by window
- `output_to_share/tables/03_sf_trajectory_data.csv` — SF ratio by window
- `output_to_share/tables/03_fio2_overall_data.csv` 
- `output_to_share/tables/03_sf_overall_data.csv` 
- `output_to_share/tables/03_hfno_spline_data.csv` — overall trajectory
- `output_to_share/tables/05_events_by_window.csv` — event counts by window
- `output_to_share/tables/06_correlation_matrix.csv` — confounder correlations

**Models (return all):**

- `output_to_share/models/07_weight_diagnostics.csv` — ESS by window ⭐
- `output_to_share/models/07_weight_distribution.csv` 
- `output_to_share/models/08_model3_exposure_effect.csv` — **primary result** ⭐
- `output_to_share/models/08_model3_unweighted_full.csv` — full unweighted coefficients
- `output_to_share/models/08_model3_weighted_full.csv` — full weighted coefficients
- `output_to_share/models/09_clustering_adjusted_results.csv` — clustering-adjusted result ⭐
- `output_to_share/models/09_icc.csv` — ICC (if applicable)
- `output_to_share/models/10_sensitivity_secondary_outcome.csv`
- `output_to_share/models/10_sensitivity_untruncated_weights.csv`
- `output_to_share/models/10_sensitivity_sf_confounder.csv`
- `output_to_share/models/10_sensitivity_spline_exposure.csv`

**Diagnostics:**

- `output_to_share/diagnostics/01_baseline_completeness.csv`
- `output_to_share/diagnostics/01_late_events.csv`
- `output_to_share/session_info.txt`

**Figures** (optional — return if flagged issues):

- `output_to_share/figures/03_hfno_over_time.png`
- `output_to_share/figures/03_fio2_over_time.png`
- `output_to_share/figures/03_sf_over_time.png`
- `output_to_share/figures/03_baseline_characteristics.png`
- `output_to_share/figures/05_events_by_window.png`
- `output_to_share/figures/06_baseline_hfno_distribution.png`
- `output_to_share/figures/06_baseline_hfno_vs_fio2.png`
- `output_to_share/figures/06_baseline_hfno_vs_sf.png`
- `output_to_share/figures/07_weight_distribution.png`
