---
title: "ROX Index as Predictor of Post-Extubation ICU Outcomes"
subtitle: "Federated Predictive Analysis — Site-Level Script"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    embed-resources: true
    fig-width: 11
    fig-height: 6
execute:
  warning: false
  message: false
  echo: true
---

# Overview

This document implements analysis examining the ROX index as a predictor of post-extubation ICU outcomes. Sites should run this document from top to bottom without modification. All outputs are saved to the `output_to_share/` directory for return to the coordinating site.

**Study Design:**

- **Predictor:** ROX index (SpO₂/FiO₂ / Respiratory Rate), measured at 4h, 8h, and 12h post-extubation using worst-case values within each window (minimum SpO₂, maximum FiO₂, maximum RR)
- **Outcomes (8 total):** Extubation success/failure at 3-day and 7-day horizons, under two definitions (standard and strict); death or reintubation at 12h and 24h, under two definitions (standard and with NIPPV/CPAP)
- **Method:** Logistic regression with cluster-robust standard errors; AUROC comparison across timepoints; 
**Required data file:** `rox_prediction.parquet` in the `output/` directory

**Required variables:**

| Variable | Description |
|---|---|
| `hospitalization_id` | Unique hospitalization identifier |
| `age` | Age at admission (years) |
| `sex` | Sex category |
| `cci` | Charlson Comorbidity Index |
| `sofa_extubation` | SOFA score in 24h prior to extubation |
| `chronic_resp_failure_hypoxia` | Prior ICD-10 J96.11 diagnosis (boolean) |
| `vasopressor_0_4h` | Vasopressor use, 0–4h post-extubation (boolean) |
| `vasopressor_4_8h` | Vasopressor use, 4–8h post-extubation (boolean) |
| `vasopressor_8_12h` | Vasopressor use, 8–12h post-extubation (boolean) |
| `rox_0_4h` | ROX index, 0–4h window (nullable) |
| `rox_4_8h` | ROX index, 4–8h window (nullable) |
| `rox_8_12h` | ROX index, 8–12h window (nullable) |
| `extubation_success_3d` | Extubation success within 3 days — no reintubation, no death |
| `extubation_success_strict_3d` | Extubation success within 3 days — no reintubation, no death, no NIPPV/CPAP |
| `extubation_success_7d` | Extubation success within 7 days — no reintubation, no death |
| `extubation_success_strict_7d` | Extubation success within 7 days — no reintubation, no death, no NIPPV/CPAP |
| `death_or_reintubation_12h` | Death or reintubation within 12 hours post-extubation |
| `death_or_reintubation_nippv_cpap_12h` | Death or reintubation or NIPPV/CPAP within 12 hours post-extubation |
| `death_or_reintubation_24h` | Death or reintubation within 24 hours post-extubation |
| `death_or_reintubation_nippv_cpap_24h` | Death or reintubation or NIPPV/CPAP within 24 hours post-extubation |
| `hospital_id` | Facility identifier (set to a constant string if single-site) |

---

# Section 0 — Setup

```{r setup}
library(tidyverse)
library(arrow)
library(gtsummary)
library(pROC)
library(lmtest)
library(sandwich)
library(survival)
library(patchwork)
library(knitr)
library(scales)
library(broom)

set.seed(20240101)

# ── Output directories ──────────────────────────────────────────────────────
dir.create("../output_to_share/rox/figures",     showWarnings = FALSE, recursive = TRUE)
dir.create("../output_to_share/rox/tables",      showWarnings = FALSE, recursive = TRUE)
dir.create("../output_to_share/rox/diagnostics", showWarnings = FALSE, recursive = TRUE)
dir.create("../output_to_share/rox/models",      showWarnings = FALSE, recursive = TRUE)

# ── Load data ────────────────────────────────────────────────────────────────
df_raw <- read_parquet("../output/rox_prediction.parquet")

# ── Global constants ─────────────────────────────────────────────────────────
# ROX thresholds from published literature
ROX_ROCA_2016        <- 4.88   # Roca et al. 2016, J Crit Care — original derivation (pneumonia/ARF)
ROX_POST_EXTUBATION  <- 8.70   # Post-extubation literature threshold (e.g., 2h/6h windows)

# Named ROX windows (timepoint labels used throughout)
rox_windows <- c("4h" = "rox_0_4h", "8h" = "rox_4_8h", "12h" = "rox_8_12h")

# Outcome definitions — used to loop over all 8 outcomes (chronological order)
# "strict" = no death, no reintubation, AND no NIPPV/CPAP (NIPPV/CPAP counts as failure)
# "non-strict" = no death, no reintubation (NIPPV/CPAP allowed as success)
outcome_defs <- tibble(
  outcome_var   = c("extubation_success_strict_12h", "extubation_success_12h",
                     "extubation_success_strict_24h", "extubation_success_24h",
                     "extubation_success_strict_3d",  "extubation_success_3d",
                     "extubation_success_strict_7d",  "extubation_success_7d"),
  outcome_label = c("Extubation Failure+NIPPV/CPAP 12h", "Extubation Failure 12h",
                     "Extubation Failure+NIPPV/CPAP 24h", "Extubation Failure 24h",
                     "Extubation Failure+NIPPV/CPAP 3d",  "Extubation Failure 3d",
                     "Extubation Failure+NIPPV/CPAP 7d",  "Extubation Failure 7d"),
  horizon       = c("12h", "12h", "24h", "24h", "3d", "3d", "7d", "7d"),
  strict        = c(TRUE, FALSE, TRUE, FALSE, TRUE, FALSE, TRUE, FALSE)
)

# Corresponding vasopressor column for each ROX window
vaso_for_window <- c("4h" = "vasopressor_0_4h", "8h" = "vasopressor_4_8h", "12h" = "vasopressor_8_12h")
# ── TEMPORARY: dummy hospital_id for single-site pilot testing ─────────────
# Remove or comment out before distributing to sites
if (!"hospital_id" %in% names(df_raw)) {
  df_raw <- df_raw %>% mutate(hospital_id = "PILOT_SITE")
  cat("NOTE: hospital_id not found — dummy variable 'PILOT_SITE' created for testing.\n")
}
cat("Data loaded:", nrow(df_raw), "rows,",
    n_distinct(df_raw$hospitalization_id), "hospitalizations\n")
cat("Sites:", n_distinct(df_raw$hospital_id), "\n")
```

---

# Section 1 — Data Quality and Missingness

```{r data-quality}
# ── 1.1 Structure ─────────────────────────────────────────────────────────
cat("Rows:", nrow(df_raw), "\n")
cat("Columns:", ncol(df_raw), "\n")
cat("Unique hospitalizations:", n_distinct(df_raw$hospitalization_id), "\n")
cat("Unique hospitals:", n_distinct(df_raw$hospital_id), "\n")

# ── 1.2 Duplicate check — expect 0 ────────────────────────────────────────
dups <- df_raw %>% group_by(hospitalization_id) %>% filter(n() > 1)
if (nrow(dups) > 0) {
  warning(nrow(dups), " duplicate hospitalization_id rows detected. Review before proceeding.")
  print(dups)
} else {
  cat("Duplicate check: PASS\n")
}

# ── 1.3 Outcome completeness ──────────────────────────────────────────────
cat("\nOutcome variable missingness:\n")
outcome_miss <- df_raw %>%
  summarise(
    across(starts_with("extubation_success"), ~ sum(is.na(.x))),
    across(c(extubation_success_strict_12h, extubation_success_12h,
                   extubation_success_strict_24h, extubation_success_24h), ~ sum(is.na(.x)))
  ) %>%
  pivot_longer(everything(), names_to = "outcome", values_to = "n_missing")
print(outcome_miss)

# ── 1.4 ROX missingness by window ─────────────────────────────────────────
rox_miss <- df_raw %>%
  summarise(
    n_total      = n(),
    rox_4h_miss  = sum(is.na(rox_0_4h)),
    rox_8h_miss  = sum(is.na(rox_4_8h)),
    rox_12h_miss = sum(is.na(rox_8_12h)),
    rox_4h_pct   = round(rox_4h_miss  / n_total * 100, 1),
    rox_8h_pct   = round(rox_8h_miss  / n_total * 100, 1),
    rox_12h_pct  = round(rox_12h_miss / n_total * 100, 1)
  )

cat("\nROX index missingness by window:\n")
print(rox_miss)

# ── 1.5 Covariate completeness ────────────────────────────────────────────
covariate_miss <- df_raw %>%
  summarise(
    n              = n(),
    age_miss       = sum(is.na(age)),
    sex_miss       = sum(is.na(sex)),
    cci_miss       = sum(is.na(cci)),
    sofa_miss      = sum(is.na(sofa_extubation)),
    chr_resp_miss  = sum(is.na(chronic_resp_failure_hypoxia)),
    vaso_4h_miss   = sum(is.na(vasopressor_0_4h)),
    vaso_8h_miss   = sum(is.na(vasopressor_4_8h)),
    vaso_12h_miss  = sum(is.na(vasopressor_8_12h))
  )

cat("\nCovariate completeness (n missing):\n")
print(covariate_miss)

# ── 1.6 ROX missingness by outcome group ─────────────────────────────────
# Flags if missingness is differential — important for complete-case validity
rox_miss_by_outcome <- df_raw %>%
  mutate(
    outcome_group = case_when(
       extubation_success_7d == TRUE  ~ "Success (7d)",
       extubation_success_7d == FALSE ~ "Failure (7d)",
       TRUE                           ~ "Unknown"
    )
  ) %>%
  group_by(outcome_group) %>%
  summarise(
    n             = n(),
    rox_4h_miss   = sum(is.na(rox_0_4h)),
    rox_8h_miss   = sum(is.na(rox_4_8h)),
    rox_12h_miss  = sum(is.na(rox_8_12h)),
    rox_4h_pct    = round(rox_4h_miss  / n * 100, 1),
    rox_8h_pct    = round(rox_8h_miss  / n * 100, 1),
    rox_12h_pct   = round(rox_12h_miss / n * 100, 1),
    .groups = "drop"
  )

cat("\nROX missingness by 7-day outcome group (flag if differential):\n")
print(rox_miss_by_outcome)

# ── 1.7 ROX missingness by site ───────────────────────────────────────────
rox_miss_by_site <- df_raw %>%
  group_by(hospital_id) %>%
  summarise(
    n             = n(),
    rox_4h_pct    = round(sum(is.na(rox_0_4h))  / n * 100, 1),
    rox_8h_pct    = round(sum(is.na(rox_4_8h))  / n * 100, 1),
    rox_12h_pct   = round(sum(is.na(rox_8_12h)) / n * 100, 1),
    .groups = "drop"
  )

cat("\nROX missingness by site (flag sites with >30% missing at any window):\n")
print(rox_miss_by_site)

problematic_sites <- rox_miss_by_site %>%
  filter(rox_4h_pct > 30 | rox_8h_pct > 30 | rox_12h_pct > 30)
if (nrow(problematic_sites) > 0) {
  warning("Sites with >30% ROX missingness at any window — results may be unreliable for these sites:")
  print(problematic_sites)
}

# ── Save diagnostics ──────────────────────────────────────────────────────
write_csv(rox_miss,            "../output_to_share/rox/diagnostics/01_rox_missingness_overall.csv")
write_csv(rox_miss_by_outcome, "../output_to_share/rox/diagnostics/01_rox_missingness_by_outcome.csv")
write_csv(rox_miss_by_site,    "../output_to_share/rox/diagnostics/01_rox_missingness_by_site.csv")
write_csv(covariate_miss,      "../output_to_share/rox/diagnostics/01_covariate_missingness.csv")
```

---

# Section 2 — Cohort Description and Table 1

```{r cohort}
# ── 2.1 Outcome distribution ──────────────────────────────────────────────
outcome_dist <- df_raw %>%
  summarise(
    n_total                    = n(),
    # 12h outcomes (strict = not allowed NIPPV/CPAP as part of failure; non-strict = does not matter whether you use NIPPV CPAP)
    n_success_strict_12h       = sum( extubation_success_strict_12h,          na.rm = TRUE),
    n_failure_strict_12h       = sum(!extubation_success_strict_12h,          na.rm = TRUE),
    n_success_12h              = sum( extubation_success_12h,                 na.rm = TRUE),
    n_failure_12h              = sum(!extubation_success_12h,                 na.rm = TRUE),
    # 24h outcomes (strict = not allowed NIPPV/CPAP as part of failure; non-strict = does not matter whether you use NIPPV CPAP)
    n_success_strict_24h       = sum( extubation_success_strict_24h,          na.rm = TRUE),
    n_failure_strict_24h       = sum(!extubation_success_strict_24h,          na.rm = TRUE),
    n_success_24h              = sum( extubation_success_24h,                 na.rm = TRUE),
    n_failure_24h              = sum(!extubation_success_24h,                 na.rm = TRUE),
    # 3-day (strict = not allowed NIPPV/CPAP as part of failure; non-strict = does not matter whether you use NIPPV CPAP)
    n_success_3d               = sum( extubation_success_3d,        na.rm = TRUE),
    n_failure_3d               = sum(!extubation_success_3d,        na.rm = TRUE),
    n_success_strict_3d        = sum( extubation_success_strict_3d, na.rm = TRUE),
    n_failure_strict_3d        = sum(!extubation_success_strict_3d, na.rm = TRUE),
    # 7-day (strict = not allowed NIPPV/CPAP as part of failure; non-strict = does not matter whether you use NIPPV CPAP)
    n_success_7d               = sum( extubation_success_7d,        na.rm = TRUE),
    n_failure_7d               = sum(!extubation_success_7d,        na.rm = TRUE),
    n_success_strict_7d        = sum( extubation_success_strict_7d, na.rm = TRUE),
    n_failure_strict_7d        = sum(!extubation_success_strict_7d, na.rm = TRUE)
  ) %>%
  pivot_longer(everything()) %>%
  mutate(
    pct = round(value / value[name == "n_total"] * 100, 1)
  )

cat("Outcome distribution:\n")
print(outcome_dist)
write_csv(outcome_dist, "../output_to_share/rox/tables/02_outcome_distribution.csv")

# ── 2.2 Prepare analysis dataset ─────────────────────────────────────────
# Add failure indicators (FALSE = success, TRUE = failure) for convenience
df <- df_raw %>%
  mutate(
    # All outcomes now use same convention (TRUE = success)
    failure_strict_12h = !extubation_success_strict_12h,
    failure_12h        = !extubation_success_12h,
    failure_strict_24h = !extubation_success_strict_24h,
    failure_24h        = !extubation_success_24h,
    failure_3d        = !extubation_success_3d,
    failure_strict_3d = !extubation_success_strict_3d,
    failure_7d        = !extubation_success_7d,
    failure_strict_7d = !extubation_success_strict_7d,
    sex               = factor(sex),
    chronic_resp_failure_hypoxia = as.integer(chronic_resp_failure_hypoxia)
  )

# ── 2.3 Table 1 — by 12h extubation outcome ─────────────────────────────
df %>%
  select(
    age, sex, cci, chronic_resp_failure_hypoxia, sofa_extubation,
    vasopressor_0_4h, rox_0_4h, rox_4_8h, rox_8_12h,
    extubation_success_strict_12h
  ) %>%
  mutate(
    outcome_group = factor(
      ifelse(extubation_success_strict_12h, "Success (12h)", "Failure (12h)"),
      levels = c("Success (12h)", "Failure (12h)")
    )
  ) %>%
  select(-extubation_success_strict_12h) %>%
  tbl_summary(
    by = outcome_group,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) | {median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = list(
      age                          ~ "Age, years",
      sex                          ~ "Sex",
      cci                          ~ "Charlson Comorbidity Index",
      chronic_resp_failure_hypoxia ~ "Prior Chronic Respiratory Failure with Hypoxia (J96.11)",
      sofa_extubation              ~ "SOFA Score at Extubation",
      vasopressor_0_4h             ~ "Vasopressor Use, 0–4h Post-Extubation",
      rox_0_4h                     ~ "ROX Index at 4h",
      rox_4_8h                     ~ "ROX Index at 8h",
      rox_8_12h                    ~ "ROX Index at 12h"
    ),
    missing = "ifany"
  ) %>%
  add_overall(last = FALSE) %>%
  bold_labels() %>%
  as_tibble() %>%
  write_csv("../output_to_share/rox/tables/02_table1_by_12h_outcome.csv")
cat("Table 1 (12h) saved.\n")

# ── 2.4 Table 1 — by 24h extubation outcome ─────────────────────────────
df %>%
  select(
    age, sex, cci, chronic_resp_failure_hypoxia, sofa_extubation,
    vasopressor_0_4h, rox_0_4h, rox_4_8h, rox_8_12h,
    extubation_success_strict_24h
  ) %>%
  mutate(
    outcome_group = factor(
      ifelse(extubation_success_strict_24h, "Success (24h)", "Failure (24h)"),
      levels = c("Success (24h)", "Failure (24h)")
    )
  ) %>%
  select(-extubation_success_strict_24h) %>%
  tbl_summary(
    by = outcome_group,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) | {median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = list(
      age                          ~ "Age, years",
      sex                          ~ "Sex",
      cci                          ~ "Charlson Comorbidity Index",
      chronic_resp_failure_hypoxia ~ "Prior Chronic Respiratory Failure with Hypoxia (J96.11)",
      sofa_extubation              ~ "SOFA Score at Extubation",
      vasopressor_0_4h             ~ "Vasopressor Use, 0–4h Post-Extubation",
      rox_0_4h                     ~ "ROX Index at 4h",
      rox_4_8h                     ~ "ROX Index at 8h",
      rox_8_12h                    ~ "ROX Index at 12h"
    ),
    missing = "ifany"
  ) %>%
  add_overall(last = FALSE) %>%
  bold_labels() %>%
  as_tibble() %>%
  write_csv("../output_to_share/rox/tables/02_table1_by_24h_outcome.csv")
cat("Table 1 (24h) saved.\n")

# ── 2.5 Table 1 — by 3-day outcome ──────────────────────────────────────
df %>%
  select(
    age, sex, cci, chronic_resp_failure_hypoxia, sofa_extubation,
    vasopressor_0_4h, rox_0_4h, rox_4_8h, rox_8_12h,
    extubation_success_3d
  ) %>%
  mutate(
    outcome_group = factor(
      ifelse(extubation_success_3d, "Success (3d)", "Failure (3d)"),
      levels = c("Success (3d)", "Failure (3d)")
    )
  ) %>%
  select(-extubation_success_3d) %>%
  tbl_summary(
    by = outcome_group,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) | {median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = list(
      age                          ~ "Age, years",
      sex                          ~ "Sex",
      cci                          ~ "Charlson Comorbidity Index",
      chronic_resp_failure_hypoxia ~ "Prior Chronic Respiratory Failure with Hypoxia (J96.11)",
      sofa_extubation              ~ "SOFA Score at Extubation",
      vasopressor_0_4h             ~ "Vasopressor Use, 0–4h Post-Extubation",
      rox_0_4h                     ~ "ROX Index at 4h",
      rox_4_8h                     ~ "ROX Index at 8h",
      rox_8_12h                    ~ "ROX Index at 12h"
    ),
    missing = "ifany"
  ) %>%
  add_overall(last = FALSE) %>%
  bold_labels() %>%
  as_tibble() %>% 
  write_csv("../output_to_share/rox/tables/02_table1_by_3d_outcome.csv")
cat("Table 1 (3-day) saved.\n")

# ── 2.6 Table 1 — by 7-day outcome ──────────────────────────────────────
df %>%
  select(
    age, sex, cci, chronic_resp_failure_hypoxia, sofa_extubation,
    vasopressor_0_4h, rox_0_4h, rox_4_8h, rox_8_12h,
    extubation_success_7d
  ) %>%
  mutate(
    outcome_group = factor(
      ifelse(extubation_success_7d, "Success (7d)", "Failure (7d)"),
      levels = c("Success (7d)", "Failure (7d)")
    )
  ) %>%
  select(-extubation_success_7d) %>%
  tbl_summary(
    by = outcome_group,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) | {median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = list(
      age                          ~ "Age, years",
      sex                          ~ "Sex",
      cci                          ~ "Charlson Comorbidity Index",
      chronic_resp_failure_hypoxia ~ "Prior Chronic Respiratory Failure with Hypoxia (J96.11)",
      sofa_extubation              ~ "SOFA Score at Extubation",
      vasopressor_0_4h             ~ "Vasopressor Use, 0–4h Post-Extubation",
      rox_0_4h                     ~ "ROX Index at 4h",
      rox_4_8h                     ~ "ROX Index at 8h",
      rox_8_12h                    ~ "ROX Index at 12h"
    ),
    missing = "ifany"
  ) %>%
  add_overall(last = FALSE) %>%
  bold_labels() %>%
  as_tibble() %>% 
  write_csv("../output_to_share/rox/tables/02_table1_by_7d_outcome.csv")
cat("Table 1 (7-day) saved.\n")

# ── 2.7 ROX distribution by outcome — descriptive ─────────────────────────
rox_by_outcome <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "window", values_to = "rox") %>%
  mutate(window = recode(window,
    "rox_0_4h" = "4h", "rox_4_8h" = "8h", "rox_8_12h" = "12h")) %>%
  pivot_longer(
    cols = c(extubation_success_strict_12h, extubation_success_12h,
             extubation_success_strict_24h, extubation_success_24h,
             extubation_success_3d, extubation_success_strict_3d,
             extubation_success_7d, extubation_success_strict_7d),
    names_to = "outcome", values_to = "outcome_success"
  ) %>%
  mutate(outcome = recode(outcome,
    "extubation_success_strict_12h" = "12h (Strict)",
    "extubation_success_12h"        = "12h",
    "extubation_success_strict_24h" = "24h (Strict)",
    "extubation_success_24h"        = "24h",
    "extubation_success_3d"         = "3d",
    "extubation_success_strict_3d"  = "3d (Strict)",
    "extubation_success_7d"         = "7d",
    "extubation_success_strict_7d"  = "7d (Strict)"),
    outcome = factor(outcome, levels = c(
      "12h (Strict)", "12h", "24h (Strict)", "24h",
      "3d", "3d (Strict)", "7d", "7d (Strict)"))
  ) %>%
  group_by(outcome, window, outcome_success) %>%
  summarise(
    n_rox      = sum(!is.na(rox)),
    n_miss_rox  = sum(is.na(rox)),
    median_rox = round(median(rox, na.rm = TRUE), 2),
    p25_rox    = round(quantile(rox, 0.25, na.rm = TRUE), 2),
    p75_rox    = round(quantile(rox, 0.75, na.rm = TRUE), 2),
    mean_rox   = round(mean(rox, na.rm = TRUE), 2),
    sd_rox     = round(sd(rox, na.rm = TRUE), 2),
    .groups = "drop"
  ) %>%
  arrange(outcome, window, desc(outcome_success)) 

cat("\nROX index by all outcome and window:\n")
print(rox_by_outcome)
write_csv(rox_by_outcome, "../output_to_share/rox/tables/02_rox_by_outcome.csv")
```

---

# Section 3 — Exploratory Visualizations

```{r viz-rox-dist-12h-density, fig.height=7}
# ── 3.1 ROX density distribution — 12h extubation outcome ──────
p_rox_dist_12h_std <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "window", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window = factor(recode(window,
      "rox_0_4h"  = "4h Post-Extubation",
      "rox_4_8h"  = "8h Post-Extubation",
      "rox_8_12h" = "12h Post-Extubation"),
      levels = c("4h Post-Extubation", "8h Post-Extubation", "12h Post-Extubation")),
    outcome_label = factor(
      ifelse(extubation_success_strict_12h, "Success (12h)", "Failure (12h)"),
      levels = c("Success (12h)", "Failure (12h)")
    )
  ) %>%
  ggplot(aes(x = rox, fill = outcome_label, color = outcome_label)) +
  geom_density(alpha = 0.35, linewidth = 0.8) +
  geom_vline(xintercept = ROX_ROCA_2016,      linetype = "dotted", color = "grey30", linewidth = 0.7) +
  geom_vline(xintercept = ROX_POST_EXTUBATION, linetype = "dashed", color = "grey30", linewidth = 0.7) +
  annotate("text", x = ROX_ROCA_2016,      y = Inf, label = "Roca ≥4.88",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  annotate("text", x = ROX_POST_EXTUBATION, y = Inf, label = "Post-ext. ≥8.7",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  scale_fill_manual(values  = c("Success (12h)" = "#27AE60", "Failure (12h)" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success (12h)" = "#27AE60", "Failure (12h)" = "#C0392B"), name = NULL) +
  facet_wrap(~window, ncol = 1) +
  labs(title    = "ROX Index Distribution by 12h Extubation Outcome",
       subtitle = "Dotted = Roca 2016 threshold (4.88); Dashed = post-extubation threshold (8.70)",
       x = "ROX Index", y = "Density") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank(),
        legend.position = "bottom")

ggsave("../output_to_share/rox/figures/03_rox_distribution_12h_std_outcome.png",
       plot = p_rox_dist_12h_std, width = 9, height = 9, dpi = 300)
p_rox_dist_12h_std
```

```{r viz-rox-dist-12h-nippv-density, fig.height=7}
# ── 3.2 ROX density distribution — 12h extubation+NIPPV/CPAP outcome ───
p_rox_dist_12h_nippv <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "window", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window = factor(recode(window,
      "rox_0_4h"  = "4h Post-Extubation",
      "rox_4_8h"  = "8h Post-Extubation",
      "rox_8_12h" = "12h Post-Extubation"),
      levels = c("4h Post-Extubation", "8h Post-Extubation", "12h Post-Extubation")),
    outcome_label = factor(
      ifelse(extubation_success_12h, "Success (12h)", "Failure+NIPPV/CPAP (12h)"),
      levels = c("Success (12h)", "Failure+NIPPV/CPAP (12h)")
    )
  ) %>%
  ggplot(aes(x = rox, fill = outcome_label, color = outcome_label)) +
  geom_density(alpha = 0.35, linewidth = 0.8) +
  geom_vline(xintercept = ROX_ROCA_2016,      linetype = "dotted", color = "grey30", linewidth = 0.7) +
  geom_vline(xintercept = ROX_POST_EXTUBATION, linetype = "dashed", color = "grey30", linewidth = 0.7) +
  annotate("text", x = ROX_ROCA_2016,      y = Inf, label = "Roca ≥4.88",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  annotate("text", x = ROX_POST_EXTUBATION, y = Inf, label = "Post-ext. ≥8.7",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  scale_fill_manual(values  = c("Success (12h)" = "#27AE60", "Failure+NIPPV/CPAP (12h)" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success (12h)" = "#27AE60", "Failure+NIPPV/CPAP (12h)" = "#C0392B"), name = NULL) +
  facet_wrap(~window, ncol = 1) +
  labs(title    = "ROX Index Distribution by 12h Extubation+NIPPV/CPAP Outcome",
       subtitle = "Dotted = Roca 2016 threshold (4.88); Dashed = post-extubation threshold (8.70)",
       x = "ROX Index", y = "Density") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank(),
        legend.position = "bottom")

ggsave("../output_to_share/rox/figures/03_rox_distribution_12h_nippv_outcome.png",
       plot = p_rox_dist_12h_nippv, width = 9, height = 9, dpi = 300)
p_rox_dist_12h_nippv
```

```{r viz-rox-dist-24h-density, fig.height=7}
# ── 3.3 ROX density distribution — 24h extubation outcome ──────
p_rox_dist_24h_std <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "window", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window = factor(recode(window,
      "rox_0_4h"  = "4h Post-Extubation",
      "rox_4_8h"  = "8h Post-Extubation",
      "rox_8_12h" = "12h Post-Extubation"),
      levels = c("4h Post-Extubation", "8h Post-Extubation", "12h Post-Extubation")),
    outcome_label = factor(
      ifelse(extubation_success_strict_24h, "Success (24h)", "Failure (24h)"),
      levels = c("Success (24h)", "Failure (24h)")
    )
  ) %>%
  ggplot(aes(x = rox, fill = outcome_label, color = outcome_label)) +
  geom_density(alpha = 0.35, linewidth = 0.8) +
  geom_vline(xintercept = ROX_ROCA_2016,      linetype = "dotted", color = "grey30", linewidth = 0.7) +
  geom_vline(xintercept = ROX_POST_EXTUBATION, linetype = "dashed", color = "grey30", linewidth = 0.7) +
  annotate("text", x = ROX_ROCA_2016,      y = Inf, label = "Roca ≥4.88",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  annotate("text", x = ROX_POST_EXTUBATION, y = Inf, label = "Post-ext. ≥8.7",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  scale_fill_manual(values  = c("Success (24h)" = "#27AE60", "Failure (24h)" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success (24h)" = "#27AE60", "Failure (24h)" = "#C0392B"), name = NULL) +
  facet_wrap(~window, ncol = 1) +
  labs(title    = "ROX Index Distribution by 24h Extubation Outcome",
       subtitle = "Dotted = Roca 2016 threshold (4.88); Dashed = post-extubation threshold (8.70)",
       x = "ROX Index", y = "Density") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank(),
        legend.position = "bottom")

ggsave("../output_to_share/rox/figures/03_rox_distribution_24h_std_outcome.png",
       plot = p_rox_dist_24h_std, width = 9, height = 9, dpi = 300)
p_rox_dist_24h_std
```

```{r viz-rox-dist-24h-nippv-density, fig.height=7}
# ── 3.4 ROX density distribution — 24h extubation+NIPPV/CPAP outcome ───
p_rox_dist_24h_nippv <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "window", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window = factor(recode(window,
      "rox_0_4h"  = "4h Post-Extubation",
      "rox_4_8h"  = "8h Post-Extubation",
      "rox_8_12h" = "12h Post-Extubation"),
      levels = c("4h Post-Extubation", "8h Post-Extubation", "12h Post-Extubation")),
    outcome_label = factor(
      ifelse(extubation_success_24h, "Success (24h)", "Failure+NIPPV/CPAP (24h)"),
      levels = c("Success (24h)", "Failure+NIPPV/CPAP (24h)")
    )
  ) %>%
  ggplot(aes(x = rox, fill = outcome_label, color = outcome_label)) +
  geom_density(alpha = 0.35, linewidth = 0.8) +
  geom_vline(xintercept = ROX_ROCA_2016,      linetype = "dotted", color = "grey30", linewidth = 0.7) +
  geom_vline(xintercept = ROX_POST_EXTUBATION, linetype = "dashed", color = "grey30", linewidth = 0.7) +
  annotate("text", x = ROX_ROCA_2016,      y = Inf, label = "Roca ≥4.88",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  annotate("text", x = ROX_POST_EXTUBATION, y = Inf, label = "Post-ext. ≥8.7",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  scale_fill_manual(values  = c("Success (24h)" = "#27AE60", "Failure+NIPPV/CPAP (24h)" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success (24h)" = "#27AE60", "Failure+NIPPV/CPAP (24h)" = "#C0392B"), name = NULL) +
  facet_wrap(~window, ncol = 1) +
  labs(title    = "ROX Index Distribution by 24h Extubation+NIPPV/CPAP Outcome",
       subtitle = "Dotted = Roca 2016 threshold (4.88); Dashed = post-extubation threshold (8.70)",
       x = "ROX Index", y = "Density") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank(),
        legend.position = "bottom")

ggsave("../output_to_share/rox/figures/03_rox_distribution_24h_nippv_outcome.png",
       plot = p_rox_dist_24h_nippv, width = 9, height = 9, dpi = 300)
p_rox_dist_24h_nippv
```

```{r viz-rox-dist-3d, fig.height=7}
# ── 3.5 ROX density distribution — 3-day outcome ─────────────────────────
p_rox_dist_3d <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "window", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window = factor(recode(window,
      "rox_0_4h"  = "4h Post-Extubation",
      "rox_4_8h"  = "8h Post-Extubation",
      "rox_8_12h" = "12h Post-Extubation"),
      levels = c("4h Post-Extubation", "8h Post-Extubation", "12h Post-Extubation")),
    outcome_label = factor(
      ifelse(extubation_success_3d, "Success (3d)", "Failure (3d)"),
      levels = c("Success (3d)", "Failure (3d)")
    )
  ) %>%
  ggplot(aes(x = rox, fill = outcome_label, color = outcome_label)) +
  geom_density(alpha = 0.35, linewidth = 0.8) +
  geom_vline(xintercept = ROX_ROCA_2016,      linetype = "dotted", color = "grey30", linewidth = 0.7) +
  geom_vline(xintercept = ROX_POST_EXTUBATION, linetype = "dashed", color = "grey30", linewidth = 0.7) +
  annotate("text", x = ROX_ROCA_2016,      y = Inf, label = "Roca ≥4.88",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  annotate("text", x = ROX_POST_EXTUBATION, y = Inf, label = "Post-ext. ≥8.7",
           vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  scale_fill_manual(values  = c("Success (3d)" = "#27AE60", "Failure (3d)" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success (3d)" = "#27AE60", "Failure (3d)" = "#C0392B"), name = NULL) +
  facet_wrap(~window, ncol = 1) +
  labs(title    = "ROX Index Distribution by 3-Day Outcome",
       subtitle = "Dotted = Roca 2016 threshold (4.88); Dashed = post-extubation threshold (8.70)",
       x = "ROX Index", y = "Density") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank(),
        legend.position = "bottom")

ggsave("../output_to_share/rox/figures/03_rox_distribution_3d_outcome.png",
       plot = p_rox_dist_3d, width = 9, height = 9, dpi = 300)
p_rox_dist_3d
```

```{r viz-rox-distributions, fig.height=7}
# ── 3.6 ROX density distribution — 7-day outcome ─────────────────────────
p_rox_dist <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "window", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window        = factor(recode(window,
      "rox_0_4h" = "4h Post-Extubation",
      "rox_4_8h" = "8h Post-Extubation",
      "rox_8_12h" = "12h Post-Extubation"),
      levels = c("4h Post-Extubation", "8h Post-Extubation", "12h Post-Extubation")),
    outcome_label = factor(
      ifelse(extubation_success_7d, "Success (7d)", "Failure (7d)"),
      levels = c("Success (7d)", "Failure (7d)")
    )
  ) %>%
  ggplot(aes(x = rox, fill = outcome_label, color = outcome_label)) +
  geom_density(alpha = 0.35, linewidth = 0.8) +
  geom_vline(xintercept = ROX_ROCA_2016,       linetype = "dotted",  color = "grey30", linewidth = 0.7) +
  geom_vline(xintercept = ROX_POST_EXTUBATION,  linetype = "dashed",  color = "grey30", linewidth = 0.7) +
  annotate("text", x = ROX_ROCA_2016,       y = Inf, label = "Roca ≥4.88",      vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  annotate("text", x = ROX_POST_EXTUBATION,  y = Inf, label = "Post-ext. ≥8.7", vjust = 1.5, hjust = -0.1, size = 3, color = "grey30") +
  scale_fill_manual(values  = c("Success (7d)" = "#27AE60", "Failure (7d)" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success (7d)" = "#27AE60", "Failure (7d)" = "#C0392B"), name = NULL) +
  facet_wrap(~window, ncol = 1) +
  labs(title    = "ROX Index Distribution by 7-Day Outcome",
       subtitle = "Dotted = Roca 2016 threshold (4.88); Dashed = post-extubation threshold (8.70)",
       x = "ROX Index", y = "Density") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"), panel.grid.minor = element_blank(),
        legend.position = "bottom")

ggsave("../output_to_share/rox/figures/03_rox_distribution_by_outcome.png",
       plot = p_rox_dist, width = 9, height = 9, dpi = 300)
p_rox_dist
```

```{r viz-rox-boxplots-12h, fig.height=9}
# ── 3.7 ROX boxplot — 12h outcomes ─────────────────────────
df_box_12h <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "rox_col", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window = factor(recode(rox_col,
      "rox_0_4h"  = "4h",
      "rox_4_8h"  = "8h",
      "rox_8_12h" = "12h"),
      levels = c("4h", "8h", "12h"))
  ) %>%
  pivot_longer(
    cols      = c(extubation_success_strict_12h, extubation_success_12h),
    names_to  = "horizon",
    values_to = "success"
  ) %>%
  mutate(
    horizon       = recode(horizon,
      "extubation_success_strict_12h"          = "Extubation Failure 12h",
      "extubation_success_12h"                 = "Extubation Failure+NIPPV/CPAP 12h"),
    horizon       = factor(horizon, levels = c("Extubation Failure 12h",
                                                "Extubation Failure+NIPPV/CPAP 12h")),
    outcome_label = factor(
      ifelse(success, "Success", "Failure"),
      levels = c("Success", "Failure")
    )
  )

p_rox_box_12h <- ggplot(df_box_12h,
                        aes(x = window, y = rox,
                            fill = outcome_label, color = outcome_label)) +
  geom_boxplot(alpha = 0.4, outlier.shape = NA, width = 0.5,
               position = position_dodge(0.7)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.7),
              size = 0.6, alpha = 0.3) +
  geom_hline(yintercept = ROX_POST_EXTUBATION, linetype = "dashed",
             color = "grey40", linewidth = 0.7) +
  scale_fill_manual(values  = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  facet_wrap(~horizon, ncol = 1) +
  labs(title    = "ROX Index at Each Post-Extubation Window — 12h Extubation Outcomes",
       subtitle = "Dashed line = post-extubation threshold (8.70)",
       x = "Time Post-Extubation", y = "ROX Index") +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        strip.text       = element_text(face = "bold", size = 12))

ggsave("../output_to_share/rox/figures/03_rox_boxplot_12h_outcomes.png",
       plot = p_rox_box_12h, width = 9, height = 10, dpi = 300)
p_rox_box_12h
```

```{r viz-rox-boxplots-24h, fig.height=9}
# ── 3.8 ROX boxplot — 24h outcomes ─────────────────────────
df_box_24h <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "rox_col", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window = factor(recode(rox_col,
      "rox_0_4h"  = "4h",
      "rox_4_8h"  = "8h",
      "rox_8_12h" = "12h"),
      levels = c("4h", "8h", "12h"))
  ) %>%
  pivot_longer(
    cols      = c(extubation_success_strict_24h, extubation_success_24h),
    names_to  = "horizon",
    values_to = "success"
  ) %>%
  mutate(
    horizon       = recode(horizon,
      "extubation_success_strict_24h"          = "Extubation Failure 24h",
      "extubation_success_24h"                 = "Extubation Failure+NIPPV/CPAP 24h"),
    horizon       = factor(horizon, levels = c("Extubation Failure 24h",
                                                "Extubation Failure+NIPPV/CPAP 24h")),
    outcome_label = factor(
      ifelse(success, "Success", "Failure"),
      levels = c("Success", "Failure")
    )
  )

p_rox_box_24h <- ggplot(df_box_24h,
                        aes(x = window, y = rox,
                            fill = outcome_label, color = outcome_label)) +
  geom_boxplot(alpha = 0.4, outlier.shape = NA, width = 0.5,
               position = position_dodge(0.7)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.7),
              size = 0.6, alpha = 0.3) +
  geom_hline(yintercept = ROX_POST_EXTUBATION, linetype = "dashed",
             color = "grey40", linewidth = 0.7) +
  scale_fill_manual(values  = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  facet_wrap(~horizon, ncol = 1) +
  labs(title    = "ROX Index at Each Post-Extubation Window — 24h Extubation Outcomes",
       subtitle = "Dashed line = post-extubation threshold (8.70)",
       x = "Time Post-Extubation", y = "ROX Index") +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        strip.text       = element_text(face = "bold", size = 12))

ggsave("../output_to_share/rox/figures/03_rox_boxplot_24h_outcomes.png",
       plot = p_rox_box_24h, width = 9, height = 10, dpi = 300)
p_rox_box_24h
```
```{r viz-rox-boxplots, fig.height=9}
# ── 3.9 ROX boxplot — 3d outcomes ─────────────────────────
df_box_3d <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "rox_col", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window = factor(recode(rox_col,
      "rox_0_4h"  = "4h",
      "rox_4_8h"  = "8h",
      "rox_8_12h" = "12h"),
      levels = c("4h", "8h", "12h"))
  ) %>%
  pivot_longer(
    cols      = c(extubation_success_3d, extubation_success_strict_3d),
    names_to  = "horizon",
    values_to = "success"
  ) %>%
  mutate(
    horizon       = recode(horizon,
      "extubation_success_3d"         = "Extubation Failure 3d",
      "extubation_success_strict_3d"  = "Extubation Failure+NIPPV/CPAP 3d"),
    horizon       = factor(horizon, levels = c("Extubation Failure 3d",
                                                "Extubation Failure+NIPPV/CPAP 3d")),
    outcome_label = factor(
      ifelse(success, "Success", "Failure"),
      levels = c("Success", "Failure")
    )
  )

p_rox_box_3d <- ggplot(df_box_3d,
                        aes(x = window, y = rox,
                            fill = outcome_label, color = outcome_label)) +
  geom_boxplot(alpha = 0.4, outlier.shape = NA, width = 0.5,
               position = position_dodge(0.7)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.7),
              size = 0.6, alpha = 0.3) +
  geom_hline(yintercept = ROX_POST_EXTUBATION, linetype = "dashed",
             color = "grey40", linewidth = 0.7) +
  scale_fill_manual(values  = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  facet_wrap(~horizon, ncol = 1) +
  labs(title    = "ROX Index at Each Post-Extubation Window — 3d Extubation Outcomes",
       subtitle = "Dashed line = post-extubation threshold (8.70)",
       x = "Time Post-Extubation", y = "ROX Index") +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        strip.text       = element_text(face = "bold", size = 12))

ggsave("../output_to_share/rox/figures/03_rox_boxplot_3d_outcomes.png",
       plot = p_rox_box_3d, width = 9, height = 10, dpi = 300)
p_rox_box_3d

```
```{r viz-rox-boxplots, fig.height=9}
# ── 3.10 ROX boxplot — 7d outcomes ─────────────────────────
df_box_7d <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "rox_col", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  mutate(
    window = factor(recode(rox_col,
      "rox_0_4h"  = "4h",
      "rox_4_8h"  = "8h",
      "rox_8_12h" = "12h"),
      levels = c("4h", "8h", "12h"))
  ) %>%
  pivot_longer(
    cols      = c(extubation_success_7d, extubation_success_strict_7d),
    names_to  = "horizon",
    values_to = "success"
  ) %>%
  mutate(
    horizon       = recode(horizon,
      "extubation_success_7d"         = "Extubation Failure 7d",
      "extubation_success_strict_7d"  = "Extubation Failure+NIPPV/CPAP 7d"),
    horizon       = factor(horizon, levels = c("Extubation Failure 7d",
                                                "Extubation Failure+NIPPV/CPAP 7d")),
    outcome_label = factor(
      ifelse(success, "Success", "Failure"),
      levels = c("Success", "Failure")
    )
  )

p_rox_box_7d <- ggplot(df_box_7d,
                        aes(x = window, y = rox,
                            fill = outcome_label, color = outcome_label)) +
  geom_boxplot(alpha = 0.4, outlier.shape = NA, width = 0.5,
               position = position_dodge(0.7)) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.7),
              size = 0.6, alpha = 0.3) +
  geom_hline(yintercept = ROX_POST_EXTUBATION, linetype = "dashed",
             color = "grey40", linewidth = 0.7) +
  scale_fill_manual(values  = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_color_manual(values = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  facet_wrap(~horizon, ncol = 1) +
  labs(title    = "ROX Index at Each Post-Extubation Window — 7d Extubation Outcomes",
       subtitle = "Dashed line = post-extubation threshold (8.70)",
       x = "Time Post-Extubation", y = "ROX Index") +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        strip.text       = element_text(face = "bold", size = 12))

ggsave("../output_to_share/rox/figures/03_rox_boxplot_7d_outcomes.png",
       plot = p_rox_box_7d, width = 9, height = 10, dpi = 300)
p_rox_box_7d
```

```{r viz-rox-trajectory-12h, fig.height=7}
# ── 3.10 Median ROX trajectory across windows — 12h outcomes ────────────
df_traj_12h <- bind_rows(
  df %>%
    pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
                 names_to = "rox_col", values_to = "rox") %>%
    filter(!is.na(rox)) %>%
    mutate(
      horizon       = "Extubation Failure 12h",
      outcome_label = factor(ifelse(extubation_success_strict_12h, "Success", "Failure"),
                             levels = c("Success", "Failure"))
    ),
  df %>%
    pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
                 names_to = "rox_col", values_to = "rox") %>%
    filter(!is.na(rox)) %>%
    mutate(
      horizon       = "Extubation Failure+NIPPV/CPAP 12h",
      outcome_label = factor(ifelse(extubation_success_12h, "Success", "Failure"),
                             levels = c("Success", "Failure"))
    )
) %>%
  mutate(
    window  = factor(recode(rox_col,
      "rox_0_4h"  = "4h",
      "rox_4_8h"  = "8h",
      "rox_8_12h" = "12h"),
      levels = c("4h", "8h", "12h")),
    horizon = factor(horizon, levels = c("Extubation Failure 12h",
                                          "Extubation Failure+NIPPV/CPAP 12h"))
  ) %>%
  group_by(horizon, window, outcome_label) %>%
  summarise(
    n      = sum(!is.na(rox)),
    median = median(rox, na.rm = TRUE),
    p25    = quantile(rox, 0.25, na.rm = TRUE),
    p75    = quantile(rox, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

p_rox_traj_12h <- ggplot(df_traj_12h,
                     aes(x = window, y = median, group = outcome_label,
                         color = outcome_label, fill = outcome_label)) +
  geom_ribbon(aes(ymin = p25, ymax = p75), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(aes(size = n), alpha = 0.85) +
  geom_hline(yintercept = ROX_POST_EXTUBATION, linetype = "dashed",
             color = "grey40", linewidth = 0.7) +
  scale_color_manual(values = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_fill_manual(values  = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_size_continuous(name = "N", range = c(2, 6)) +
  facet_wrap(~horizon, ncol = 1) +
  labs(title    = "Median ROX Index Over Post-Extubation Windows — 12h Extubation Outcomes",
       subtitle = "Median ± IQR ribbon | Dashed line = post-extubation threshold (8.70)",
       x = "Time Post-Extubation", y = "ROX Index (Median)") +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        strip.text       = element_text(face = "bold", size = 12))

ggsave("../output_to_share/rox/figures/03_rox_trajectory_12h_outcome.png",
       plot = p_rox_traj_12h, width = 9, height = 10, dpi = 300)
write_csv(df_traj_12h, "../output_to_share/rox/tables/03_rox_trajectory_12h_data.csv")

p_rox_traj_12h
```

```{r viz-rox-trajectory-24h, fig.height=7}
# ── 3.11 Median ROX trajectory across windows — 24h outcomes ────────────
df_traj_24h <- bind_rows(
  df %>%
    pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
                 names_to = "rox_col", values_to = "rox") %>%
    filter(!is.na(rox)) %>%
    mutate(
      horizon       = "Extubation Failure 24h",
      outcome_label = factor(ifelse(extubation_success_strict_24h, "Success", "Failure"),
                             levels = c("Success", "Failure"))
    ),
  df %>%
    pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
                 names_to = "rox_col", values_to = "rox") %>%
    filter(!is.na(rox)) %>%
    mutate(
      horizon       = "Extubation Failure+NIPPV/CPAP 24h",
      outcome_label = factor(ifelse(extubation_success_24h, "Success", "Failure"),
                             levels = c("Success", "Failure"))
    )
) %>%
  mutate(
    window  = factor(recode(rox_col,
      "rox_0_4h"  = "4h",
      "rox_4_8h"  = "8h",
      "rox_8_12h" = "12h"),
      levels = c("4h", "8h", "12h")),
    horizon = factor(horizon, levels = c("Extubation Failure 24h",
                                          "Extubation Failure+NIPPV/CPAP 24h"))
  ) %>%
  group_by(horizon, window, outcome_label) %>%
  summarise(
    n      = sum(!is.na(rox)),
    median = median(rox, na.rm = TRUE),
    p25    = quantile(rox, 0.25, na.rm = TRUE),
    p75    = quantile(rox, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

p_rox_traj_24h <- ggplot(df_traj_24h,
                     aes(x = window, y = median, group = outcome_label,
                         color = outcome_label, fill = outcome_label)) +
  geom_ribbon(aes(ymin = p25, ymax = p75), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(aes(size = n), alpha = 0.85) +
  geom_hline(yintercept = ROX_POST_EXTUBATION, linetype = "dashed",
             color = "grey40", linewidth = 0.7) +
  scale_color_manual(values = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_fill_manual(values  = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_size_continuous(name = "N", range = c(2, 6)) +
  facet_wrap(~horizon, ncol = 1) +
  labs(title    = "Median ROX Index Over Post-Extubation Windows — 24h Extubation Outcomes",
       subtitle = "Median ± IQR ribbon | Dashed line = post-extubation threshold (8.70)",
       x = "Time Post-Extubation", y = "ROX Index (Median)") +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        strip.text       = element_text(face = "bold", size = 12))

ggsave("../output_to_share/rox/figures/03_rox_trajectory_24h_outcome.png",
       plot = p_rox_traj_24h, width = 9, height = 10, dpi = 300)
write_csv(df_traj_24h, "../output_to_share/rox/tables/03_rox_trajectory_24h_data.csv")

p_rox_traj_24h
```

```{r viz-rox-trajectory-3d, fig.height=7}
# ── 3.12a Median ROX trajectory across windows — 3-day ───────────────────
df_traj_3d <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "rox_col", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  pivot_longer(
    cols      = c(extubation_success_3d, extubation_success_strict_3d),
    names_to  = "horizon",
    values_to = "success"
  ) %>%
  mutate(
    window  = factor(recode(rox_col,
      "rox_0_4h"  = "4h",
      "rox_4_8h"  = "8h",
      "rox_8_12h" = "12h"),
      levels = c("4h", "8h", "12h")),
    horizon = recode(horizon,
      "extubation_success_3d"        = "Extubation Failure 3d",
      "extubation_success_strict_3d" = "Extubation Failure+NIPPV/CPAP 3d"),
    horizon = factor(horizon, levels = c("Extubation Failure 3d",
                                          "Extubation Failure+NIPPV/CPAP 3d")),
    outcome_label = factor(ifelse(success, "Success", "Failure"),
                           levels = c("Success", "Failure"))
  ) %>%
  group_by(horizon, window, outcome_label) %>%
  summarise(
    n      = sum(!is.na(rox)),
    median = median(rox, na.rm = TRUE),
    p25    = quantile(rox, 0.25, na.rm = TRUE),
    p75    = quantile(rox, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

p_rox_traj_3d <- ggplot(df_traj_3d,
                     aes(x = window, y = median, group = outcome_label,
                         color = outcome_label, fill = outcome_label)) +
  geom_ribbon(aes(ymin = p25, ymax = p75), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(aes(size = n), alpha = 0.85) +
  geom_hline(yintercept = ROX_POST_EXTUBATION, linetype = "dashed",
             color = "grey40", linewidth = 0.7) +
  scale_color_manual(values = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_fill_manual(values  = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_size_continuous(name = "N", range = c(2, 6)) +
  facet_wrap(~horizon, ncol = 1) +
  labs(title    = "Median ROX Index Over Post-Extubation Windows — 3d Extubation Outcomes",
       subtitle = "Median ± IQR ribbon | Dashed line = post-extubation threshold (8.70)",
       x = "Time Post-Extubation", y = "ROX Index (Median)") +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        strip.text       = element_text(face = "bold", size = 12))

ggsave("../output_to_share/rox/figures/03_rox_trajectory_3d_outcome.png",
       plot = p_rox_traj_3d, width = 9, height = 10, dpi = 300)
write_csv(df_traj_3d, "../output_to_share/rox/tables/03_rox_trajectory_3d_data.csv")
p_rox_traj_3d
```
```{r viz-rox-trajectory-7d, fig.height=7}
# ── 3.12b Median ROX trajectory across windows — 7-day ───────────────────
df_traj_7d <- df %>%
  pivot_longer(cols = c(rox_0_4h, rox_4_8h, rox_8_12h),
               names_to = "rox_col", values_to = "rox") %>%
  filter(!is.na(rox)) %>%
  pivot_longer(
    cols      = c(extubation_success_7d, extubation_success_strict_7d),
    names_to  = "horizon",
    values_to = "success"
  ) %>%
  mutate(
    window  = factor(recode(rox_col,
      "rox_0_4h"  = "4h",
      "rox_4_8h"  = "8h",
      "rox_8_12h" = "12h"),
      levels = c("4h", "8h", "12h")),
    horizon = recode(horizon,
      "extubation_success_7d"        = "Extubation Failure 7d",
      "extubation_success_strict_7d" = "Extubation Failure+NIPPV/CPAP 7d"),
    horizon = factor(horizon, levels = c("Extubation Failure 7d",
                                          "Extubation Failure+NIPPV/CPAP 7d")),
    outcome_label = factor(ifelse(success, "Success", "Failure"),
                           levels = c("Success", "Failure"))
  ) %>%
  group_by(horizon, window, outcome_label) %>%
  summarise(
    n      = sum(!is.na(rox)),
    median = median(rox, na.rm = TRUE),
    p25    = quantile(rox, 0.25, na.rm = TRUE),
    p75    = quantile(rox, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

p_rox_traj_7d <- ggplot(df_traj_7d,
                     aes(x = window, y = median, group = outcome_label,
                         color = outcome_label, fill = outcome_label)) +
  geom_ribbon(aes(ymin = p25, ymax = p75), alpha = 0.15, color = NA) +
  geom_line(linewidth = 1.1) +
  geom_point(aes(size = n), alpha = 0.85) +
  geom_hline(yintercept = ROX_POST_EXTUBATION, linetype = "dashed",
             color = "grey40", linewidth = 0.7) +
  scale_color_manual(values = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_fill_manual(values  = c("Success" = "#27AE60", "Failure" = "#C0392B"), name = NULL) +
  scale_size_continuous(name = "N", range = c(2, 6)) +
  facet_wrap(~horizon, ncol = 1) +
  labs(title    = "Median ROX Index Over Post-Extubation Windows — 7d Extubation Outcomes",
       subtitle = "Median ± IQR ribbon | Dashed line = post-extubation threshold (8.70)",
       x = "Time Post-Extubation", y = "ROX Index (Median)") +
  theme_minimal(base_size = 13) +
  theme(plot.title       = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position  = "bottom",
        strip.text       = element_text(face = "bold", size = 12))

ggsave("../output_to_share/rox/figures/03_rox_trajectory_7d_outcome.png",
       plot = p_rox_traj_7d, width = 9, height = 10, dpi = 300)
write_csv(df_traj_7d, "../output_to_share/rox/tables/03_rox_trajectory_7d_data.csv")
p_rox_traj_7d
```


```{r viz-rox-covariate, fig.height=8}
# ── 3.14 Baseline covariate distributions by 3-day and 7-day outcome ───────
# Mirrors the HFNO file's Section 3.4 baseline characteristics panel

make_covariate_panel <- function(outcome_col, outcome_label_success,
                                  outcome_label_failure, colors, title_suffix) {
  df %>%
    mutate(
      outcome_group = factor(
        ifelse(.data[[outcome_col]], outcome_label_success, outcome_label_failure),
        levels = c(outcome_label_success, outcome_label_failure)
      )
    ) %>%
    select(outcome_group, age, sofa_extubation, cci,
           rox_0_4h, rox_4_8h, rox_8_12h,
           vasopressor_0_4h, chronic_resp_failure_hypoxia) %>%
    mutate(
      vasopressor_0_4h             = as.numeric(vasopressor_0_4h),
      chronic_resp_failure_hypoxia = as.numeric(chronic_resp_failure_hypoxia)
    ) %>%
    pivot_longer(cols = -outcome_group, names_to = "variable", values_to = "value") %>%
    mutate(var_label = recode(variable,
      "age"                          = "Age (years)",
      "sofa_extubation"              = "SOFA at Extubation",
      "cci"                          = "Charlson Index",
      "rox_0_4h"                     = "ROX at 4h",
      "rox_4_8h"                     = "ROX at 8h",
      "rox_8_12h"                    = "ROX at 12h",
      "vasopressor_0_4h"             = "Vasopressors (0–4h)",
      "chronic_resp_failure_hypoxia" = "Prior Chronic Resp. Failure"
    )) %>%
    filter(!is.na(value)) %>%
    ggplot(aes(x = outcome_group, y = value,
               fill = outcome_group, color = outcome_group)) +
    geom_boxplot(alpha = 0.4, outlier.shape = NA, width = 0.5) +
    geom_jitter(width = 0.15, size = 0.8, alpha = 0.4) +
    scale_fill_manual(values  = colors, guide = "none") +
    scale_color_manual(values = colors, guide = "none") +
    facet_wrap(~var_label, scales = "free_y", ncol = 4) +
    labs(title = paste("Baseline Characteristics by", title_suffix),
         x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(plot.title  = element_text(face = "bold"),
          strip.text  = element_text(face = "bold"),
          axis.text.x = element_text(angle = 20, hjust = 1),
          panel.grid.minor = element_blank())
}

p_cov_12h_strict <- make_covariate_panel(
  outcome_col            = "extubation_success_strict_12h",
  outcome_label_success  = "Success (12h)",
  outcome_label_failure  = "Failure (12h)",
  colors                 = c("Success (12h)" = "#27AE60", "Failure (12h)" = "#C0392B"),
  title_suffix           = "12h Extubation Failure")

p_cov_12h <- make_covariate_panel(
  outcome_col            = "extubation_success_12h",
  outcome_label_success  = "Success (12h)",
  outcome_label_failure  = "Failure+NIPPV/CPAP (12h)",
  colors                 = c("Success (12h)" = "#27AE60", "Failure+NIPPV/CPAP (12h)" = "#C0392B"),
  title_suffix           = "12h Extubation Failure+NIPPV/CPAP")

p_cov_24h_strict <- make_covariate_panel(
  outcome_col            = "extubation_success_strict_24h",
  outcome_label_success  = "Success (24h)",
  outcome_label_failure  = "Failure (24h)",
  colors                 = c("Success (24h)" = "#27AE60", "Failure (24h)" = "#C0392B"),
  title_suffix           = "24h Extubation Failure")

p_cov_24h <- make_covariate_panel(
  outcome_col            = "extubation_success_24h",
  outcome_label_success  = "Success (24h)",
  outcome_label_failure  = "Failure+NIPPV/CPAP (24h)",
  colors                 = c("Success (24h)" = "#27AE60", "Failure+NIPPV/CPAP (24h)" = "#C0392B"),
  title_suffix           = "24h Extubation Failure+NIPPV/CPAP")

p_cov_3d_strict <- make_covariate_panel(
  outcome_col            = "extubation_success_strict_3d",
  outcome_label_success  = "Success (3d)",
  outcome_label_failure  = "Failure+NIPPV/CPAP (3d)",
  colors                 = c("Success (3d)" = "#27AE60", "Failure+NIPPV/CPAP (3d)" = "#C0392B"),
  title_suffix           = "3d Extubation Failure+NIPPV/CPAP")

p_cov_3d <- make_covariate_panel(
  outcome_col            = "extubation_success_3d",
  outcome_label_success  = "Success (3d)",
  outcome_label_failure  = "Failure (3d)",
  colors                 = c("Success (3d)" = "#27AE60", "Failure (3d)" = "#C0392B"),
  title_suffix           = "3d Extubation Failure")

p_cov_7d_strict <- make_covariate_panel(
  outcome_col            = "extubation_success_strict_7d",
  outcome_label_success  = "Success (7d)",
  outcome_label_failure  = "Failure+NIPPV/CPAP (7d)",
  colors                 = c("Success (7d)" = "#27AE60", "Failure+NIPPV/CPAP (7d)" = "#C0392B"),
  title_suffix           = "7d Extubation Failure+NIPPV/CPAP")

p_cov_7d <- make_covariate_panel(
  outcome_col            = "extubation_success_7d",
  outcome_label_success  = "Success (7d)",
  outcome_label_failure  = "Failure (7d)",
  colors                 = c("Success (7d)" = "#27AE60", "Failure (7d)" = "#C0392B"),
  title_suffix           = "7d Extubation Failure")

ggsave("../output_to_share/rox/figures/03_covariates_12h_strict_outcome.png",
       plot = p_cov_12h_strict, width = 13, height = 8, dpi = 300)
ggsave("../output_to_share/rox/figures/03_covariates_12h_outcome.png",
       plot = p_cov_12h, width = 13, height = 8, dpi = 300)
ggsave("../output_to_share/rox/figures/03_covariates_24h_strict_outcome.png",
       plot = p_cov_24h_strict, width = 13, height = 8, dpi = 300)
ggsave("../output_to_share/rox/figures/03_covariates_24h_outcome.png",
       plot = p_cov_24h, width = 13, height = 8, dpi = 300)
ggsave("../output_to_share/rox/figures/03_covariates_3d_outcome.png",
       plot = p_cov_3d, width = 13, height = 8, dpi = 300)
ggsave("../output_to_share/rox/figures/03_covariates_3d_strict_outcome.png",
       plot = p_cov_3d_strict, width = 13, height = 8, dpi = 300)
ggsave("../output_to_share/rox/figures/03_covariates_7d_outcome.png",
       plot = p_cov_7d, width = 13, height = 8, dpi = 300)
ggsave("../output_to_share/rox/figures/03_covariates_7d_strict_outcome.png",
       plot = p_cov_7d_strict, width = 13, height = 8, dpi = 300)

p_cov_12h_strict
p_cov_12h
p_cov_24h_strict
p_cov_24h
p_cov_3d
p_cov_3d_strict
p_cov_7d
p_cov_7d_strict
```

---

# Section 4 — AUROC Analysis

ROX index (continuous) is evaluated as a predictor of each of the 8 outcomes at each of the 3 timepoints. This yields 24 ROC curves total (8 outcomes × 3 timepoints). DeLong's test compares AUROCs across timepoints within each outcome.

**Note on complete-case analysis:** All ROC and regression analyses use only patients with non-missing ROX values at the relevant window. Missingness counts are reported in Section 1. If a site has substantial or differential missingness, interpret results with caution and flag for the coordinating site.

```{r auroc}
# ── 4.1 Compute AUROCs for all outcome × timepoint combinations ───────────

# Failure indicators (chronological order: 12h, 24h, 3d, 7d)
failure_vars <- c(
  "Extubation Failure 12h"            = "failure_12h",
  "Extubation Failure+NIPPV/CPAP 12h" = "failure_strict_12h",
  "Extubation Failure 24h"            = "failure_24h",
  "Extubation Failure+NIPPV/CPAP 24h" = "failure_strict_24h",
  "Extubation Failure 3d"             = "failure_3d",
  "Extubation Failure+NIPPV/CPAP 3d"  = "failure_strict_3d",
  "Extubation Failure 7d"             = "failure_7d",
  "Extubation Failure+NIPPV/CPAP 7d"  = "failure_strict_7d"
)

rox_col_labels <- c(
  "rox_0_4h"  = "4h",
  "rox_4_8h"  = "8h",
  "rox_8_12h" = "12h"
)

# Store ROC objects for later plotting and DeLong comparison
roc_objects <- list()

auroc_results <- map_dfr(names(failure_vars), function(out_label) {
  fv <- failure_vars[[out_label]]
  map_dfr(names(rox_col_labels), function(rox_col) {
    tp_label <- rox_col_labels[[rox_col]]
    dat      <- df %>% filter(!is.na(.data[[rox_col]]))
    n_total  <- nrow(dat)
    n_fail   <- sum(dat[[fv]], na.rm = TRUE)
    n_succ   <- n_total - n_fail

    if (n_fail < 5 | n_succ < 5) {
      warning("Insufficient events for ROC at ", out_label, " / ", tp_label,
              " (", n_fail, " failures, ", n_succ, " successes) — skipping.")
      return(NULL)
    }

    roc_obj <- roc(dat[[fv]], dat[[rox_col]], direction = "auto",
                   quiet = TRUE, ci = TRUE)

    roc_objects[[paste0(out_label, "__", tp_label)]] <<- roc_obj

    tibble(
      outcome       = out_label,
      timepoint     = tp_label,
      n_total       = n_total,
      n_failure     = n_fail,
      n_success     = n_succ,
      auc           = round(as.numeric(roc_obj$auc), 3),
      auc_lower     = round(as.numeric(roc_obj$ci[1]), 3),
      auc_upper     = round(as.numeric(roc_obj$ci[3]), 3),
      # ">" = higher ROX associated with more failure (inverse of Roca 2016 expectation)
      # "<" = higher ROX associated with less failure (expected per published literature)
      roc_direction = roc_obj$direction
    )
  })
})

# ── Flag unexpected direction ─────────────────────────────────────────────
unexpected_dir <- auroc_results %>% filter(roc_direction == ">")
if (nrow(unexpected_dir) > 0) {
  message("NOTE: ", nrow(unexpected_dir), " ROC(s) have direction='>' (higher ROX = more failure). ",
          "This may reflect FiO2 titration in response to deterioration in operational EHR data. ",
          "AUCs are reported correctly by pROC. Flag for coordinating site.")
  print(unexpected_dir %>% select(outcome, timepoint, auc, roc_direction))
}

kable(auroc_results,
      caption = "AUROC by Outcome and ROX Timepoint (95% CI via DeLong)",
      col.names = c("Outcome", "Timepoint", "N Total", "N Failure",
                    "N Success", "AUROC", "95% CI Lower", "95% CI Upper", "ROC Direction"))
write_csv(auroc_results, "../output_to_share/rox/models/04_auroc_results.csv")
```

```{r delong}
# ── 4.2 DeLong's test — compare timepoints within each outcome ────────────
delong_results <- map_dfr(names(failure_vars), function(out_label) {
  pairs <- list(c("4h", "8h"), c("4h", "12h"), c("8h", "12h"))
  map_dfr(pairs, function(pair) {
    key1 <- paste0(out_label, "__", pair[1])
    key2 <- paste0(out_label, "__", pair[2])
    if (!key1 %in% names(roc_objects) | !key2 %in% names(roc_objects)) return(NULL)
    tryCatch({
      test <- roc.test(roc_objects[[key1]], roc_objects[[key2]], method = "delong")
      tibble(
        outcome         = out_label,
        timepoint_1     = pair[1],
        timepoint_2     = pair[2],
        auc_1           = round(as.numeric(roc_objects[[key1]]$auc), 3),
        auc_2           = round(as.numeric(roc_objects[[key2]]$auc), 3),
        delong_p        = round(test$p.value, 4)
      )
    }, error = function(e) {
      cat("DeLong test failed for", out_label, pair[1], "vs", pair[2], ":", conditionMessage(e), "\n")
      NULL
    })
  })
})

kable(delong_results,
      caption = "DeLong's Test — Pairwise Comparison of AUROCs Across Timepoints")
write_csv(delong_results, "../output_to_share/rox/models/04_delong_tests.csv")
```

```{r roc-plots, fig.height=16}
# ── 4.3 ROC curves — 4 panels (one per outcome), 3 lines per panel ─────────
timepoint_colors <- c("4h" = "#2980B9", "8h" = "#E67E22", "12h" = "#8E44AD")

roc_plot_list <- map(names(failure_vars), function(out_label) {
  fv <- failure_vars[[out_label]]

  # Build ggplot data from all 3 ROC objects for this outcome
  roc_df <- map_dfr(names(rox_col_labels), function(rox_col) {
    tp_label <- rox_col_labels[[rox_col]]
    key      <- paste0(out_label, "__", tp_label)
    if (!key %in% names(roc_objects)) return(NULL)
    roc_obj <- roc_objects[[key]]
    tibble(
      specificity = roc_obj$specificities,
      sensitivity = roc_obj$sensitivities,
      timepoint   = tp_label,
      auc_label   = sprintf("%s  AUC=%.3f", tp_label,
                            as.numeric(roc_obj$auc))
    )
  })

  if (nrow(roc_df) == 0) return(NULL)

  ggplot(roc_df, aes(x = 1 - specificity, y = sensitivity,
                     color = timepoint, linetype = timepoint)) +
    geom_line(linewidth = 1.0) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted", color = "grey60") +
    scale_color_manual(
      values = timepoint_colors,
      labels = roc_df %>% distinct(timepoint, auc_label) %>% deframe(),
      name   = NULL
    ) +
    scale_linetype_manual(
      values = c("4h" = "solid", "8h" = "dashed", "12h" = "dotdash"),
      labels = roc_df %>% distinct(timepoint, auc_label) %>% deframe(),
      name   = NULL
    ) +
    labs(title = out_label,
         x = "1 – Specificity (FPR)", y = "Sensitivity (TPR)") +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(face = "bold", size = 11),
          legend.position = "bottom",
          panel.grid.minor = element_blank())
})

# Remove NULLs and combine
roc_plot_list <- compact(roc_plot_list)

if (length(roc_plot_list) > 0) {
  p_roc_combined <- wrap_plots(roc_plot_list, ncol = 2) +
    plot_annotation(
      title    = "ROC Curves: ROX Index Predicting Post-Extubation Outcomes",
      subtitle = "Each panel = one outcome definition; lines = ROX timepoint",
      theme    = theme(plot.title    = element_text(face = "bold", size = 14),
                       plot.subtitle = element_text(size = 11))
    )

  ggsave("../output_to_share/rox/figures/04_roc_curves_all_outcomes.png",
         plot = p_roc_combined, width = 14, height = 20, dpi = 300)
  p_roc_combined
} else {
  cat("No ROC curves could be generated — check event counts.\n")
}
```

---

# Section 5 — Optimal Threshold Derivation

Empirical thresholds derived using the Youden index (maximizing sensitivity + specificity – 1) are reported alongside the published Roca 2016 threshold (≥4.88) and post-extubation literature threshold (≥8.70).

```{r thresholds}
# ── 5.1 Youden-optimal thresholds for all 12 combinations ────────────────
threshold_results <- map_dfr(names(failure_vars), function(out_label) {
  fv <- failure_vars[[out_label]]
  map_dfr(names(rox_col_labels), function(rox_col) {
    tp_label <- rox_col_labels[[rox_col]]
    key      <- paste0(out_label, "__", tp_label)
    if (!key %in% names(roc_objects)) return(NULL)

    roc_obj <- roc_objects[[key]]
    best    <- coords(roc_obj, "best", best.method = "youden",
                      ret = c("threshold", "sensitivity", "specificity",
                              "ppv", "npv", "accuracy"))

    # Also evaluate at published thresholds
    eval_thresh <- function(thr) {
      dat    <- df %>% filter(!is.na(.data[[rox_col]]))
      pred   <- as.integer(dat[[rox_col]] >= thr)
      actual <- as.integer(dat[[fv]])
      tp <- sum(pred == 1 & actual == 1)
      fp <- sum(pred == 1 & actual == 0)
      tn <- sum(pred == 0 & actual == 0)
      fn <- sum(pred == 0 & actual == 1)
      list(
        sens = round(tp / (tp + fn), 3),
        spec = round(tn / (tn + fp), 3),
        ppv  = round(tp / (tp + fp), 3),
        npv  = round(tn / (tn + fn), 3)
      )
    }

    roca_perf    <- eval_thresh(ROX_ROCA_2016)
    postex_perf  <- eval_thresh(ROX_POST_EXTUBATION)

    tibble(
      outcome               = out_label,
      timepoint             = tp_label,
      # Youden
      youden_threshold      = round(best$threshold, 2),
      youden_sensitivity    = round(best$sensitivity, 3),
      youden_specificity    = round(best$specificity, 3),
      youden_ppv            = round(best$ppv, 3),
      youden_npv            = round(best$npv, 3),
      # Roca 2016
      roca_threshold        = ROX_ROCA_2016,
      roca_sensitivity      = roca_perf$sens,
      roca_specificity      = roca_perf$spec,
      roca_ppv              = roca_perf$ppv,
      roca_npv              = roca_perf$npv,
      # Post-extubation
      postex_threshold      = ROX_POST_EXTUBATION,
      postex_sensitivity    = postex_perf$sens,
      postex_specificity    = postex_perf$spec,
      postex_ppv            = postex_perf$ppv,
      postex_npv            = postex_perf$npv
    )
  })
})

# Print in wide format for readability
kable(
  threshold_results %>%
    select(outcome, timepoint,
           youden_threshold, youden_sensitivity, youden_specificity, youden_ppv, youden_npv),
  caption = "5.1 Youden-Optimal Thresholds — Sensitivity, Specificity, PPV, NPV"
)

kable(
  threshold_results %>%
    select(outcome, timepoint,
           roca_sensitivity, roca_specificity, roca_ppv, roca_npv,
           postex_sensitivity, postex_specificity, postex_ppv, postex_npv),
  caption = "5.2 Published Threshold Performance (Roca ≥4.88 and Post-Extubation ≥8.70)"
)

write_csv(threshold_results, "../output_to_share/rox/models/05_threshold_performance.csv")
```

---

# Section 6 — Adjusted Logistic Regression

For each of the 24 outcome × timepoint combinations, two logistic regression models are fitted:

- **Model A — Continuous ROX:** ROX as a continuous predictor alongside covariates
- **Model B — Dichotomized ROX:** ROX ≥ empirical Youden threshold (binary) alongside covariates

**Covariates in all models:** age, sex, CCI, chronic respiratory failure with hypoxia, SOFA at extubation, window-matched vasopressor use

**Clustering:** Site fixed effects added if >1 hospital site; cluster-robust standard errors applied in all multi-site runs using the sandwich estimator.

```{r regression-setup}
# ── 6.0 Detect clustering scenario ───────────────────────────────────────
n_sites        <- n_distinct(df$hospital_id)
multi_site     <- n_sites > 1
use_site_fe    <- multi_site     # site fixed effects if multi-site

cat("Number of sites:", n_sites, "\n")
cat("Multi-site run:", multi_site, "\n")
cat("Site fixed effects:", use_site_fe, "\n")
cat("Cluster-robust SEs:", multi_site, "\n")

if (!multi_site) {
  cat("Single-site run: standard logistic regression without clustering adjustments.\n")
}
```

```{r regression-models}
# ── Helper: fit logistic model and extract OR with 95% CI ─────────────────
fit_logistic <- function(outcome_var, rox_col, rox_binary_col,
                         vasopressor_col, data, use_fe, multi_site) {

  # Base formula covariates
  base_covars <- "age + sex + cci + chronic_resp_failure_hypoxia + sofa_extubation"
  vaso_term   <- vasopressor_col
  fe_term     <- if (use_fe) "+ hospital_id" else ""

  # Model A: continuous ROX
  fmla_cont <- as.formula(
    paste0(outcome_var, " ~ ", rox_col, " + ", base_covars,
           " + ", vaso_term, fe_term)
  )

  # Model B: dichotomized ROX
  fmla_bin <- as.formula(
    paste0(outcome_var, " ~ ", rox_binary_col, " + ", base_covars,
           " + ", vaso_term, fe_term)
  )

  fit_and_extract <- function(fmla, model_type) {
    tryCatch({
      fit <- glm(fmla, data = data, family = binomial(link = "logit"),
                 na.action = na.omit)

      # Cluster-robust SEs if multi-site
      if (multi_site) {
        vcov_robust <- vcovCL(fit, cluster = ~hospital_id)
        coef_robust <- coeftest(fit, vcov = vcov_robust)
        results <- tidy(coef_robust) %>%
          filter(grepl(rox_col, term) | grepl(rox_binary_col, term)) %>%
          mutate(
            or      = round(exp(estimate), 3),
            or_low  = round(exp(estimate - 1.96 * std.error), 3),
            or_high = round(exp(estimate + 1.96 * std.error), 3),
            p_value = round(p.value, 4),
            se_type = "cluster-robust"
          )
      } else {
        results <- tidy(fit) %>%
          filter(grepl(rox_col, term) | grepl(rox_binary_col, term)) %>%
          mutate(
            or      = round(exp(estimate), 3),
            or_low  = round(exp(estimate - 1.96 * std.error), 3),
            or_high = round(exp(estimate + 1.96 * std.error), 3),
            p_value = round(p.value, 4),
            se_type = "standard"
          )
      }

      results %>%
        mutate(
          model_type   = model_type,
          n_obs        = nobs(fit),
          n_outcome    = sum(data[[outcome_var]], na.rm = TRUE)
        ) %>%
        select(model_type, term, or, or_low, or_high, p_value, se_type, n_obs, n_outcome)

    }, error = function(e) {
      cat("Model failed:", conditionMessage(e), "\n")
      NULL
    })
  }

  bind_rows(
    fit_and_extract(fmla_cont, "Continuous ROX"),
    fit_and_extract(fmla_bin,  "Dichotomized ROX (Youden)")
  )
}

# ── 6.1 Prepare binary ROX columns (using Youden threshold per combo) ─────
# We need the Youden thresholds computed in Section 5
df_with_binary <- df

for (i in seq_len(nrow(threshold_results))) {
  rox_col   <- unname(rox_windows[threshold_results$timepoint[i]])
  youden_t  <- threshold_results$youden_threshold[i]
  new_col   <- paste0("rox_binary_",
                      gsub(" ", "_", gsub("[^0-9A-Za-z]", "_", threshold_results$outcome[i])),
                      "_", threshold_results$timepoint[i])
  df_with_binary[[new_col]] <- as.integer(df_with_binary[[rox_col]] >= youden_t)
}

# ── 6.2 Run all 24 × 2 models ─────────────────────────────────────────────
regression_results <- map_dfr(names(failure_vars), function(out_label) {
  fv    <- failure_vars[[out_label]]
  map_dfr(names(rox_col_labels), function(rox_col) {
    tp_label    <- rox_col_labels[[rox_col]]
    vaso_col    <- vaso_for_window[[tp_label]]
    binary_col  <- paste0("rox_binary_",
                          gsub(" ", "_", gsub("[^0-9A-Za-z]", "_", out_label)),
                          "_", tp_label)

    dat <- df_with_binary %>% filter(!is.na(.data[[rox_col]]))

    result <- fit_logistic(
      outcome_var    = fv,
      rox_col        = rox_col,
      rox_binary_col = binary_col,
      vasopressor_col = vaso_col,
      data           = dat,
      use_fe         = use_site_fe,
      multi_site     = multi_site
    )

    if (!is.null(result)) {
      result %>% mutate(outcome = out_label, timepoint = tp_label, .before = 1)
    }
  })
})

kable(regression_results,
      caption = "Logistic Regression — OR for ROX Index on Post-Extubation Outcomes (Continuous and Dichotomized)",
      col.names = c("Outcome", "Timepoint", "Model", "Term", "OR", "OR 95% CI Low",
                    "OR 95% CI High", "p-value", "SE Type", "N Obs", "N Outcome"))

write_csv(regression_results, "../output_to_share/rox/models/06_regression_results.csv")
```




---

# Section 8 — Site-Stratified AUROC Summary

For federated synthesis, AUROCs are computed separately per site to allow the coordinating site to assess between-site heterogeneity.

```{r site-auroc}
if (multi_site) {

  site_auroc_results <- map_dfr(unique(df$hospital_id), function(site) {
    df_site <- df %>% filter(hospital_id == site)
    map_dfr(names(failure_vars), function(out_label) {
      fv <- failure_vars[[out_label]]
      map_dfr(names(rox_col_labels), function(rox_col) {
        tp_label <- rox_col_labels[[rox_col]]
        dat      <- df_site %>% filter(!is.na(.data[[rox_col]]))
        n_fail   <- sum(dat[[fv]], na.rm = TRUE)
        n_succ   <- nrow(dat) - n_fail

        if (n_fail < 5 | n_succ < 5) {
          return(tibble(hospital_id = site, outcome = out_label, timepoint = tp_label,
                        n = nrow(dat), n_failure = n_fail, auc = NA_real_,
                        note = "Insufficient events"))
        }

        tryCatch({
          roc_site <- roc(dat[[fv]], dat[[rox_col]], direction = "auto", quiet = TRUE, ci = TRUE)
          tibble(
            hospital_id = site,
            outcome     = out_label,
            timepoint   = tp_label,
            n           = nrow(dat),
            n_failure   = n_fail,
            auc         = round(as.numeric(roc_site$auc), 3),
            auc_lower   = round(as.numeric(roc_site$ci[1]), 3),
            auc_upper   = round(as.numeric(roc_site$ci[3]), 3),
            note        = "OK"
          )
        }, error = function(e) {
          tibble(hospital_id = site, outcome = out_label, timepoint = tp_label,
                 n = nrow(dat), n_failure = n_fail, auc = NA_real_,
                 note = conditionMessage(e))
        })
      })
    })
  })

  kable(site_auroc_results,
        caption = "Site-Stratified AUROCs — Return to Coordinating Site for Pooled Analysis")
  write_csv(site_auroc_results, "../output_to_share/rox/models/08_site_stratified_auroc.csv")

} else {
  cat("Single-site run — site-stratified AUROC not applicable.\n")
  write_csv(tibble(note = "Single-site run — site-stratified AUROC not generated"),
            "../output_to_share/rox/models/08_site_stratified_auroc.csv")
}
```

---

# Session Info

```{r session-info}
writeLines(capture.output(sessionInfo()), "../output_to_share/rox/session_info.txt")
sessionInfo()
```

---

# Output File Summary

The following files are saved to `output_to_share/rox/` for return to the coordinating site. **Return all files.**

**Diagnostics:**

- `diagnostics/01_rox_missingness_overall.csv` — overall ROX missingness by window ⭐
- `diagnostics/01_rox_missingness_by_outcome.csv` — differential missingness check ⭐
- `diagnostics/01_rox_missingness_by_site.csv` — per-site missingness ⭐
- `diagnostics/01_covariate_missingness.csv`

**Tables:**

- `tables/02_outcome_distribution.csv` — cohort outcome counts (all 8 outcomes)
- `tables/02_table1_by_7d_outcome.csv` — Table 1 by 7-day outcome ⭐
- `tables/02_table1_by_3d_outcome.csv` — Table 1 by 3-day outcome
- `tables/02_table1_by_12h_outcome.csv` — Table 1 by 12h extubation outcome ⭐
- `tables/02_table1_by_24h_outcome.csv` — Table 1 by 24h extubation outcome ⭐
- `tables/02_rox_by_outcome.csv` — ROX descriptives by outcome and window
- `tables/03_rox_trajectory_data.csv` — trajectory data (3d/7d) for coordinating site
- `tables/03_rox_trajectory_12h_data.csv` — trajectory data (12h outcomes) for coordinating site
- `tables/03_rox_trajectory_24h_data.csv` — trajectory data (24h outcomes) for coordinating site
- `tables/03_rox_trajectory_composite_data.csv` — trajectory data (12h/24h composites) for coordinating site

**Models:**

- `models/04_auroc_results.csv` — AUROCs (all 24 combinations: 8 outcomes × 3 timepoints) ⭐
- `models/04_delong_tests.csv` — pairwise DeLong tests ⭐
- `models/05_threshold_performance.csv` — Youden + published threshold performance ⭐
- `models/06_regression_results.csv` — OR for ROX (continuous + dichotomized, all 8 outcomes) ⭐
- `models/07_logrank_results.csv` — log-rank test results
- `models/08_site_stratified_auroc.csv` — per-site AUROCs (all 8 outcomes) ⭐

**Figures** (return if flagged or if requested by coordinating site):

- `figures/03_rox_distribution_by_outcome.png` — density, 7-day outcome
- `figures/03_rox_distribution_3d_outcome.png` — density, 3-day outcome
- `figures/03_rox_distribution_24h_std_outcome.png` — density, 24h extubation outcome ⭐
- `figures/03_rox_distribution_24h_nippv_outcome.png` — density, 24h extubation+NIPPV/CPAP outcome ⭐
- `figures/03_rox_distribution_12h_std_outcome.png` — density, 12h extubation outcome ⭐
- `figures/03_rox_distribution_12h_nippv_outcome.png` — density, 12h extubation+NIPPV/CPAP outcome ⭐
- `figures/03_rox_boxplot_both_outcomes.png` — boxplots, 3-day and 7-day faceted ⭐
- `figures/03_rox_boxplot_12h_outcomes.png` — boxplots, 12h outcomes (standard and +NIPPV/CPAP) ⭐
- `figures/03_rox_boxplot_24h_outcomes.png` — boxplots, 24h outcomes (standard and +NIPPV/CPAP) ⭐
- `figures/03_rox_boxplot_composite_outcomes.png` — boxplots, 12h and 24h extubation outcomes ⭐
- `figures/03_rox_boxplot_composite_nippv_outcomes.png` — boxplots, 12h and 24h extubation+NIPPV/CPAP outcomes ⭐
- `figures/03_rox_trajectory_by_outcome.png` — median ROX over windows, 3d/7d horizons ⭐
- `figures/03_rox_trajectory_12h_outcome.png` — median ROX over windows, 12h outcomes ⭐
- `figures/03_rox_trajectory_24h_outcome.png` — median ROX over windows, 24h outcomes ⭐
- `figures/03_rox_trajectory_composite_outcome.png` — median ROX over windows, 12h/24h composites ⭐
- `figures/03_covariates_7d_outcome.png` — baseline characteristics panel, 7-day
- `figures/03_covariates_3d_outcome.png` — baseline characteristics panel, 3-day
- `figures/03_covariates_12h_outcome.png` — baseline characteristics panel, 12h extubation outcome
- `figures/03_covariates_12h_nippv_outcome.png` — baseline characteristics panel, 12h extubation+NIPPV/CPAP outcome
- `figures/03_covariates_24h_outcome.png` — baseline characteristics panel, 24h extubation outcome
- `figures/03_covariates_24h_nippv_outcome.png` — baseline characteristics panel, 24h extubation+NIPPV/CPAP outcome
- `figures/04_roc_curves_all_outcomes.png` — ROC curves for all 8 outcomes ⭐
- `figures/06_forest_continuous_rox.png` ⭐
- `figures/07_km_7d_standard.png` ⭐
- `figures/07_km_3d_standard.png`

⭐ = Priority files for coordinating site synthesis
